---
title: "Effective climate action can only be global in scope"
subtitle: "Insights from a Core-Periphery SOCIO-ECO-MRIO-SFC Model"
author:
  - name: "Oriol Vallès Codina"
    affiliation: "Johns Hopkins University, Net Zero Industrial Policy Lab, Department of Political Science"
    email: 
      - "oriolvallescodina@gmail.com"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc_float: true
    self-contained: true
    page-layout: full
    css: fullwidth.css
execute:
  echo: false
---

# Initialization

The initial values and variables are loaded from the file stored in variable `identif` in a different structure as before.

```{r packages, warning = F, message = F}
library(tidyverse)
library(igraph)
library(qgraph)
library(visNetwork)
library(scales)
```

```{r init, warning = FALSE, message = FALSE}
project_root <- rprojroot::find_root(
  rprojroot::is_rstudio_project | rprojroot::is_git_root
)

variable.table <<- read.csv(
  file.path(project_root, "data", "Variable_Definitions.csv")
)

# Analysis and Plotting Utils
source(file.path(project_root, "functions", "utils.R"))

identif <- file.path(project_root, "data", "initial_state_2026.xlsx")
initial <- load.new.init(identif)

source(file.path(project_root, "functions", "run_model_2026.R"))
source(file.path(project_root, "functions", "MVP_model_2026.R"))
```

# Baseline 

```{r simulation, warning = F, message = F}
# baseline <- run.model(
#   initial,
#   mvp.model,
#   log_file = file.path(project_root, "logs", "baseline_run.log"),
#   log_append = FALSE,
#   print_final_state = FALSE
# )

# saveRDS(baseline, file = file.path(project_root, 'data', 'baseline_feb10.RDS'))
baseline <- readRDS(
  file = file.path(project_root, 'data', 'baseline_feb10.RDS')
)

```

## Calibration

Target values are loaded in format long (`target.set`) and wide (`target.table.0`). Units are divided by 10,000 for convenience.

```{r calibration}
#### LOAD TARGET VALUES ####

target.list <- load.target(file.path(
  project_root,
  'data',
  'target_values.xlsx'
))
target.table <- target.list$wide
target.set <- target.list$long
target.vars <- target.set$var.label
target.set$target <- baseline$simulation[
  target.set$var,
  baseline$initial$pars['nPeriods', 'value']
]
target.set$ratio <- round(target.set$target / target.set$value, 2)

compute.target(baseline)
```

## Visualization

```{r variable_selection} 
#| fig-width: 7
#| fig-height: 4

# Variable Selection
selected.list <- list(
  Macroeconomic = c('c', 'va', 'go', 'id', 'gdef', 'cab', 'tb'),
  Social = c('n', 'nf', 'shp', 'shw'),
  Debt.and.Wealth = c('lh', 'lf', 'v', 'k', 'b_s'),
  Ecological = c('x_mat', 'land', 'mat', 'rec', 'emis', 'wa', 'water')
)


plot.selected.by_group(baseline, selected.list)
```

# Scenarios to Analyze

| Domain      | Shift              | From | To | Mechanism                  | Conserved Quantity            |
|-------------|--------------------|------|----|----------------------------|-------------------------------|
| Household Demand   | Diet         | 7    | 8  | Final demand reallocation  | Total food expenditure        |
| Household Demand   | Energy  | 31   | 32 | Final demand reallocation  | Energy service demand         |
| Production  | Wood   | 11   | 12 | Technical coefficient change | Total wood material input   |
| Production  | Pulp   | 13   | 14 | Technical coefficient change | Paper output                |
| Production  | Plastics | 17 | 18 | Technical coefficient change | Plastic product output      |
| Production  | Metal  | 26   | 27 | Technical coefficient change | Metal product output        |
| Production  | Glass  | 21   | 22 | Technical coefficient change | Glass product output        |
| Production  | Cement | 24   | 25 | Technical coefficient change | Cementitious output         |
| Production  | Energy  | 31   | 32 | Technical coefficient change | Energy service to production|
| Production  | Construction | 36   | 27 | Technical coefficient change | Construction output         |

# Scenario analysis (automated)

```{r scenario.table}
# ---- Scenario registry ----
scenarios_tbl <- tibble::tribble(
  ~shock , ~Domain            , ~Shift         , ~From , ~To , ~Mechanism                     , ~Conserved_Quantity            ,
   1L    , "Household Demand" , "Diet"         ,  7L   ,  8L , "Final demand reallocation"    , "Total food expenditure"       ,
   2L    , "Household Demand" , "Energy"       , 31L   , 32L , "Final demand reallocation"    , "Energy service demand"        ,
   3L    , "Production"       , "Wood"         , 11L   , 12L , "Technical coefficient change" , "Total wood material input"    ,
   4L    , "Production"       , "Pulp"         , 13L   , 14L , "Technical coefficient change" , "Paper output"                 ,
   5L    , "Production"       , "Plastics"     , 17L   , 18L , "Technical coefficient change" , "Plastic product output"       ,
   6L    , "Production"       , "Metal"        , 26L   , 27L , "Technical coefficient change" , "Metal product output"         ,
   7L    , "Production"       , "Glass"        , 21L   , 22L , "Technical coefficient change" , "Glass product output"         ,
   8L    , "Production"       , "Cement"       , 24L   , 25L , "Technical coefficient change" , "Cementitious output"          ,
   9L    , "Production"       , "Energy"       , 31L   , 32L , "Technical coefficient change" , "Energy service to production" ,
  10L    , "Production"       , "Construction" , 36L   , 27L , "Technical coefficient change" , "Construction output"
) %>%
  mutate(
    display_name = paste0("Scenario ", shock, ": ", Domain, " - ", Shift),
    # Parameterize these as you implement new shocks.
    shock_vars = case_when(
      shock == 1L ~ "beta-7,beta-8",
      shock == 2L ~ "beta-31,beta-32",
      TRUE ~ NA_character_
    ),
    shock_sectors = case_when(
      shock == 1L ~ "7,8",
      shock == 2L ~ "31,32",
      TRUE ~ NA_character_
    ),
    # Where the scenario run is stored (RDS). You can adopt a consistent naming scheme.
    rds_file = file.path(
      project_root,
      "data",
      paste0("shock_", shock, "_run.RDS")
    ),
    # Where the scenario summary table is stored.
    table_file = file.path(
      project_root,
      "data",
      "shock_tables",
      paste0("shock_", shock, "_table.csv")
    )
  )
```

```{r preparation_scenarios}
# ============================================================
# Scenario pipeline: registry -> compute/cache -> render
# Saves per-shock RDS artifacts under data/
# ============================================================

# -----------------------------
# User controls
# -----------------------------
run_if_missing <- FALSE # TRUE = allow running missing scenarios during render
cache_long_df <- TRUE # TRUE = cache shock.long.new() output per shock (can be large)

# -----------------------------
# Common time-period parameters
# -----------------------------
t.shock <- initial$pars["t.shock", "value"]
t0 <- 55
tf <- initial$pars["nPeriods", "value"]

# -----------------------------
# Cache locations (under data/)
# -----------------------------
data_dir <- file.path(project_root, "data")

dir_runs <- file.path(data_dir, "shock_runs")
dir_tables <- file.path(data_dir, "shock_tables")
dir_long <- file.path(data_dir, "shock_long")
dir_meta <- file.path(data_dir, "scenario_cache")

dir.create(dir_runs, showWarnings = FALSE, recursive = TRUE)
dir.create(dir_tables, showWarnings = FALSE, recursive = TRUE)
dir.create(dir_meta, showWarnings = FALSE, recursive = TRUE)
if (isTRUE(cache_long_df)) {
  dir.create(dir_long, showWarnings = FALSE, recursive = TRUE)
}

manifest_file <- file.path(dir_meta, "scenario_manifest.rds")
tables_file <- file.path(dir_meta, "scenario_tables_all.rds")

# -----------------------------
# In-memory containers (for later chunks)
# -----------------------------
scenario_runs <- list()
scenario_tables <- list()
scenario_long <- list() # only if cache_long_df = TRUE (still used during render)

# -----------------------------
# Helper: build canonical file paths from shock id
# -----------------------------
shock_paths <- function(shock) {
  shock <- as.integer(shock)
  list(
    run_rds = file.path(dir_runs, paste0("shock_", shock, "_run.rds")),
    table_rds = file.path(dir_tables, paste0("shock_", shock, "_table.rds")),
    long_rds = file.path(dir_long, paste0("shock_", shock, "_long.rds"))
  )
}
```

```{r compute_scenarios}
# ============================================================
# (1) Compute layer: load/run + compute cached derivatives
# ============================================================
compute_scenario <- function(s) {
  stopifnot(NROW(s) == 1L)

  shock <- as.integer(s$shock[[1]])
  p <- shock_paths(shock)

  # ----- Load or run model -----
  shock.run <- NULL
  if (file.exists(p$run_rds)) {
    shock.run <- readRDS(p$run_rds)
  } else if (isTRUE(run_if_missing)) {
    initial_shock <- initial
    initial_shock$pars["shock", "value"] <- shock

    shock.run <- run.model(
      initial_shock,
      mvp.model,
      log_file = file.path(
        project_root,
        "logs",
        paste0("shock_", shock, "_run.log")
      ),
      log_append = FALSE,
      print_final_state = FALSE
    )
    saveRDS(shock.run, p$run_rds)
  } else {
    stop(
      "Missing cached run for shock = ",
      shock,
      ". Expected: ",
      p$run_rds,
      "\nSet run_if_missing <- TRUE to run during render, or precompute the cache."
    )
  }

  # ----- Summary table (cached) -----
  shock.table <- NULL
  if (file.exists(p$table_rds)) {
    shock.table <- readRDS(p$table_rds)
  } else {
    shock.table <- shock.summary(
      baseline,
      shock.run = shock.run,
      t_ = c(t.shock + 1, t.shock + 7, t.shock + 20),
      t_names = c("Immediate", "Short.Term", "Long.Term")
    )
    saveRDS(shock.table, p$table_rds)
  }

  # ----- Long df (optional cached) -----
  shock.long <- NULL
  if (isTRUE(cache_long_df)) {
    if (file.exists(p$long_rds)) {
      shock.long <- readRDS(p$long_rds)
    } else {
      shock.long <- shock.long.new(baseline, shock.run)
      saveRDS(shock.long, p$long_rds)
    }
  } else {
    shock.long <- shock.long.new(baseline, shock.run)
  }

  list(
    shock = shock,
    paths = p,
    run = shock.run,
    table = shock.table,
    long = shock.long
  )
}
```

```{r render_scenarios}
# ============================================================
# (2) Render layer: headings + plots + side effects
# ============================================================
render_scenario <- function(s, obj) {
  stopifnot(NROW(s) == 1L)

  shock <- obj$shock
  shock.title <- s$display_name[[1]]

  cat("\n\n# ", shock.title, "\n\n", sep = "")

  # Self-auditing registry display
  cat(
    "- **Domain:** ",
    s$Domain[[1]],
    "\n",
    "- **Mechanism:** ",
    s$Mechanism[[1]],
    "\n",
    "- **From → To:** ",
    s$From[[1]],
    " → ",
    s$To[[1]],
    "\n",
    "- **Conserved quantity:** ",
    s$Conserved_Quantity[[1]],
    "\n\n",
    sep = ""
  )

  # ---- Visualize shocked variables ----
  cat("## Visualize shocked variables\n\n")

  has_vars <- !is.na(s$shock_vars[[1]]) && nzchar(s$shock_vars[[1]])
  has_secs <- !is.na(s$shock_sectors[[1]]) && nzchar(s$shock_sectors[[1]])

  if (isTRUE(has_vars) && isTRUE(has_secs)) {
    shock.vars <- trimws(strsplit(s$shock_vars[[1]], ",", fixed = TRUE)[[1]])
    shock.sectors <- as.integer(trimws(strsplit(
      s$shock_sectors[[1]],
      ",",
      fixed = TRUE
    )[[1]]))

    print(view.shock(
      obj$run,
      shock.vars = shock.vars,
      shock.sectors = shock.sectors,
      shock.title = shock.title
    ))
  } else {
    cat("*Pending parameterization for this scenario.*\n\n")
  }

  # ---- Visualize impacts ----
  cat("## Visualize impacts\n\n")
  df_long <- obj$long

  for (i in seq_along(selected.list)) {
    if ("gdef" %in% selected.list[[i]]) {
      print(view.vars(
        data = df_long,
        viz.vars = selected.list[[i]],
        var.label = names(selected.list)[i],
        shock.title = shock.title
      ))
    } else {
      print(view.scaled.vars(
        data = df_long,
        viz.vars = selected.list[[i]],
        var.label = names(selected.list)[i],
        shock.title = shock.title
      ))
    }
  }

  # ---- Selected variables plot ----
  cat("## Visualize selected variables\n\n")
  selected <- c("n", "c", "shw", "va", "cab", "nf", "gdef", "tb", "go")

  p.selected.vars <- dplyr::filter(df_long, variable %in% selected) %>%
    ggplot(aes(x = time, y = value, color = area, linetype = scenario)) +
    facet_wrap(~name, scales = "free") +
    geom_line() +
    geom_vline(
      xintercept = initial$pars["t.shock", "value"],
      linetype = "dashed",
      linewidth = .4
    ) +
    labs(
      title = shock.title,
      subtitle = "Selected Indicators. Vertical dashed line indicates shock time"
    ) +
    theme(
      plot.title = element_text(face = "bold"),
      plot.subtitle = element_text(face = "italic", size = 10),
      legend.title = element_text(face = "bold")
    )

  print(p.selected.vars)

  # ---- Summary visualization (existing script) ----
  cat("## Summary visualization\n\n")

  # Filenames expected by your script
  shock_table_1_filename <- file.path(
    project_root,
    "figures",
    paste0("shock_", shock, "_table_1.pdf")
  )
  shock_table_2_filename <- file.path(
    project_root,
    "figures",
    paste0("shock_", shock, "_table_2.pdf")
  )

  # Avoid locked binding for `df` by using a unique global name
  shock_df <<- obj$table
  source(file.path(project_root, "R", "scenario_analysis_2026.R"))
  rm(shock_df, envir = .GlobalEnv)

  invisible(NULL)
}
```

```{r driver_scenarios}
#| fig-width: 8
#| fig-height: 4

# ============================================================
# (3) Driver: compute everything once, then render
# ============================================================

scenarios <- seq_len(nrow(scenarios_tbl))
scenarios <- 1:2

# Compute/cache phase
for (k in scenarios) {
  s <- scenarios_tbl[k, , drop = FALSE]
  obj <- compute_scenario(s)

  scenario_runs[[as.character(obj$shock)]] <- obj$run
  scenario_tables[[as.character(obj$shock)]] <- obj$table
  scenario_long[[as.character(obj$shock)]] <- obj$long
}

# Persist manifest + aggregated tables for downstream use
saveRDS(scenarios_tbl, manifest_file)
saveRDS(scenario_tables, tables_file)

# Render phase
for (k in scenarios) {
  s <- scenarios_tbl[k, , drop = FALSE]
  shock <- as.integer(s$shock[[1]])
  key <- as.character(shock)

  obj <- list(
    shock = shock,
    run = scenario_runs[[key]],
    table = scenario_tables[[key]],
    long = scenario_long[[key]]
  )

  render_scenario(s, obj)
}
```

# Scenario Comparison

```{r comparison} 
#| fig-width: 10
#| fig-height: 21
#| fig-cap: "Comparison of Diet and Energy Transition Scenarios"

selected.list <- list(
  Macroeconomic = c('c', 'va', 'go', 'id', 'gdef', 'cab', 'tb'),
  Social = c('n', 'nf', 'shp', 'shw'),
  Debt.and.Wealth = c('lh', 'lf', 'v', 'k', 'b_s'),
  Ecological = c('x_mat', 'land', 'mat', 'rec', 'emis', 'wa', 'water')
)

lookup <- stack(selected.list) |>
  dplyr::rename(
    Variable = values,
    Name = ind
  )

scenario.tables <- setNames(
  lapply(scenarios_tbl$table_file[scenarios], read.csv),
  c("diet", "energy")
) %>%
  bind_rows(.id = "scenario") %>%
  mutate(
    scenario = factor(
      scenario,
      levels = c("diet", "energy"),
      labels = c(
        'Household Demand\nDiet Shift',
        'Household Demand\nEnergy Consumption Shift'
      )
    )
  )

df <- scenario.tables |>
  dplyr::select(-Name) |> # drop old category column
  dplyr::left_join(lookup, by = "Variable") %>%
  filter(!is.na(Name))

source(file.path(project_root, "R", "scenario_comparison_2026.R"))
p.comparison
```

# Key Structural Findings from Scenario Comparison

The scenario analysis reveals a clear macro-ecological success combined with persistent social and distributional tensions. The green transition improves environmental indicators and external balances but generates labour market contraction and rising inequality, particularly over time.

## Macroeconomic Stabilization

Across both Core and Periphery, the green transition produces systematic macroeconomic improvements, especially beyond the immediate adjustment phase.

- Current account balance and trade balance improve in all time horizons  
- Government deficit narrows significantly in the short and long term  
- Value added and gross output recover after short-term adjustment costs  
- The Core captures stronger and earlier gains than the Periphery  

**Time dynamics:**  
Short-term disruption → Medium-term stabilization → Long-term consolidation  

These patterns are consistent with a structural reallocation process rather than a prolonged recessionary adjustment.

---

## Distributional Shift Toward Capital

The transition generates a persistent and intensifying redistribution from labour to capital.

| Indicator | Immediate | Short Term | Long Term | Interpretation |
|----------|-----------|------------|-----------|----------------|
| Wage share of income | ↓ | ↓↓ | ↓↓ | Labour loses income share |
| Profit share of income | ↑ | ↑↑ | ↑↑ | Capital captures transition rents |
| Worker disposable income | ↓ | ↓ | ↓ | Real labour income compression |
| Capitalist disposable income | ↑ | ↑↑ | ↑↑ | Income polarization widens |

The shift is not temporary; it strengthens over time, indicating a structural, capital-biased transition under existing institutional settings.

---

## Labour Market Contraction

Employment effects are consistently negative, with no clear recovery in the long term.

| Variable | Core | Periphery | Dynamics |
|----------|------|-----------|----------|
| Total employment | ↓ | slight ↓ | Persistent decline |
| Female employment | ↓↓ | slight ↓ | Gendered impact |
| Real consumption | ↓ | flat/slight ↓ | Demand compression |

The results suggest a labour-saving restructuring rather than a short-term displacement followed by reabsorption. The stronger fall in female employment points to sectoral reallocation away from feminized sectors.

---

## Core–Periphery Asymmetry

The macroeconomic benefits of the transition are unevenly distributed:

- The Core experiences stronger improvements in trade balances and current account positions  
- Financial variables and wealth indicators suggest capital re-accumulation in the Core  
- The Periphery benefits less from value-added gains  

This pattern indicates upstream green upgrading concentrated in core economies, with limited peripheral value capture.

---

## Strong and Immediate Ecological Gains

Environmental improvements are large, immediate, and sustained across all horizons.

| Indicator | Pattern |
|----------|---------|
| CO₂ emissions | Large decline |
| Material extraction | Decline |
| Land use | Decline |
| Waste | Decline |
| Water use | Decline |
| Recycling | Increase |

Unlike the social variables, ecological gains are front-loaded rather than delayed.

---

## Time Horizon Summary

| Dimension | Immediate | Short Term | Long Term |
|-----------|-----------|------------|-----------|
| Ecology | Large gains | Maintained | Maintained |
| Macroeconomic stability | Mixed | Improves | Consolidates |
| Employment | Falls | Falls further | No recovery |
| Inequality | Rises | Rises more | Structural shift |
| Core–Periphery gap | Limited | Widening | Entrenched |

---

## Overall Interpretation

The modeled green transition is:

**Ecologically effective**  
**Macroeconomically stabilizing**  
**Socially regressive under current institutions**  

Without complementary policies such as:

- Wage protection and collective bargaining support  
- Employment guarantees or green job programs  
- Labour-absorbing industrial policy  
- Gender-sensitive transition measures  

the transition path risks becoming politically destabilizing despite ecological success.

---

## Executive Takeaway

- The green transition delivers rapid ecological improvements and medium-term macroeconomic stabilization, particularly in external balances and fiscal positions. 
- However, these gains are accompanied by persistent labour displacement, declining wage shares, rising profit shares, and falling worker incomes, with especially strong negative effects on female employment. 
- Inequality widens over time, and the Core captures a disproportionate share of the macroeconomic benefits relative to the Periphery.
- Without strong redistributive and labour-market institutions, the transition appears environmentally effective but socially and politically fragile.

