---
title: "Scenario Analysis"
subtitle: "Insights from a Core-Periphery SOCIO-ECO-MRIO-SFC Model"
author:
  - name: "Oriol Vallès Codina"
    affiliation: "Johns Hopkins University, Net Zero Industrial Policy Lab, Department of Political Science"
    email:
      - "oriolvallescodina@gmail.com"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc_float: true
    self-contained: true
    page-layout: full
    css: fullwidth.css
execute:
  echo: false
  working-dir: ..
---

# Initialization

This file is the **execution layer**: it runs scenarios from `scenarios.csv`, writes logs, caches `.RDS` outputs, and (optionally) produces tidy long-format outputs for downstream analysis.

```{r packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(knitr)
```

```{r init, warning=FALSE, message=FALSE}
# -----------------------------
# Robust project root detection
# -----------------------------
qmd_in <- knitr::current_input()

if (!is.null(qmd_in) && nzchar(qmd_in)) {
  qmd_path <- normalizePath(qmd_in, winslash = "/")
  root <- normalizePath(file.path(dirname(qmd_path), ".."), winslash = "/")
} else {
  wd0 <- normalizePath(getwd(), winslash = "/")
  root <- if (basename(wd0) == "qmd") normalizePath("..", winslash = "/") else wd0
}

dir_data <- file.path(root, "data")

model_dir <- file.path(root, "model")

# Workspace dirs (shared across all scenario files)
workspace_dir <- file.path(root, "output", "scenarios")
dir_runs  <- file.path(workspace_dir, "shock_runs")
dir_tables <- file.path(workspace_dir, "shock_tables")
dir_logs  <- file.path(workspace_dir, "logs")
dir_long  <- file.path(workspace_dir, "shock_long")
dir_cache <- file.path(workspace_dir, "scenario_cache")
dir_rds   <- file.path(workspace_dir, "rds_files")

invisible(lapply(
  list(workspace_dir, dir_runs, dir_tables, dir_logs, dir_long, dir_cache, dir_rds),
  dir.create,
  showWarnings = FALSE,
  recursive = TRUE
))

# Inputs
initial_filename <- file.path(dir_data, "initial_state_2026.xlsx")
stopifnot(file.exists(initial_filename))

baseline_filename <- file.path(workspace_dir, "baseline_2026.RDS")
scenario_filename <- file.path(workspace_dir, "scenarios.csv")
stopifnot(file.exists(baseline_filename))
stopifnot(file.exists(scenario_filename))

sc <- read.csv(scenario_filename, stringsAsFactors = FALSE)

# Core model bootstrapping
source(file.path(model_dir, "bootstrap_2026.R"))
initial <- load.init(initial_filename)

get_shock_filename <- function(n) file.path(dir_runs, paste0("shock_", n, "_run.RDS"))
get_log_filename   <- function(n) file.path(dir_logs, paste0("shock_", n, "_run.log"))
```

# Execution API

The pipeline expects a function `run_one_scenario(s)` in scope. If your `bootstrap_2026.R` already defines it, this block does nothing. Otherwise, a conservative default wrapper is provided.

```{r run_one_scenario, warning=FALSE, message=FALSE}
if (!exists("run_one_scenario", mode = "function")) {

  run_one_scenario <- function(s) {
    s <- as.integer(s)

    sc_row <- sc[sc$shock == s, , drop = FALSE]
    if (nrow(sc_row) != 1) {
      stop("run_one_scenario: shock id not unique or not found: ", s)
    }

    # --- Apply scenario to initial state ---
    # You must provide ONE of the following in bootstrap_2026.R:
    #   (a) apply_scenario(initial, sc, s) -> returns a modified 'initial' object, OR
    #   (b) make_shocked_initial(initial, sc_row) -> returns a modified 'initial' object.
    if (exists("apply_scenario", mode = "function")) {
      initial_shock <- apply_scenario(initial, sc = sc, shock = s)
    } else if (exists("make_shocked_initial", mode = "function")) {
      initial_shock <- make_shocked_initial(initial, sc_row = sc_row)
    } else {
      stop(
        "run_one_scenario: no scenario-applier found. Define apply_scenario() or make_shocked_initial() in model/bootstrap_2026.R."
      )
    }

    # --- Run model ---
    out <- run.model(
      initial_shock,
      mvp.model,
      log_file = get_log_filename(s),
      log_append = FALSE,
      print_final_state = FALSE
    )

    saveRDS(out, file = get_shock_filename(s))
    out
  }
}
```

# Run scenarios

```{r run_scenarios, warning=FALSE, message=FALSE}
# Choose which shocks to run:
shock_ids <- sort(unique(as.integer(sc$shock)))

# Skip shocks already cached
shock_ids <- shock_ids[!file.exists(vapply(shock_ids, get_shock_filename, character(1)))]

cat("Shocks to run:", paste(shock_ids, collapse = ", "), "\n")

# If your bootstrap defines compute_scenario(), prefer it (it may also write tables/long outputs).
if (exists("compute_scenario", mode = "function")) {
  invisible(lapply(shock_ids, function(s) {
    compute_scenario(
      shock = s,
      initial = initial,
      baseline_filename = baseline_filename,
      sc = sc,
      dir_runs = dir_runs,
      dir_tables = dir_tables,
      dir_logs = dir_logs,
      dir_long = dir_long,
      dir_cache = dir_cache
    )
  }))
} else {
  invisible(lapply(shock_ids, run_one_scenario))
}
```
