---
title: "Effective climate action can only be global in scope"
subtitle: "Insights from a Core-Periphery SOCIO-ECO-MRIO-SFC Model"
author:
  - name: "Oriol Vallès Codina"
    affiliation: "Johns Hopkins University, Net Zero Industrial Policy Lab, Department of Political Science"
    email: 
      - "oriolvallescodina@gmail.com"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc_float: true
    self-contained: true
    page-layout: full
    css: fullwidth.css
execute:
  echo: false
---

# Initialization

The initial values and variables are loaded from the file stored in variable `identif` in a different structure as before.

```{r packages, warning = F, message = F}
library(tidyverse)
library(igraph)
library(qgraph)
library(visNetwork)
library(scales)
```

```{r init, warning = FALSE, message = FALSE}
project_root <- rprojroot::find_root(
  rprojroot::is_rstudio_project | rprojroot::is_git_root
)

variable.table <<- read.csv(
  file.path(project_root, "data", "Variable_Definitions.csv")
)

source(file.path(project_root, "functions", "auxiliary_flexible_nov25.R"))

identif <- file.path(project_root, "data", "initial_state_2026.xlsx")
initial <- load.new.init(identif)

source(file.path(project_root, "functions", "run_model_2026.R"))
source(file.path(project_root, "functions", "MVP_model_2026.R"))
```

# Baseline 

```{r simulation, warning = F, message = F}
# baseline <- run.model(
#   initial,
#   mvp.model,
#   log_file = file.path(project_root, "logs", "baseline_run.log"),
#   log_append = FALSE,
#   print_final_state = FALSE
# )

# saveRDS(baseline, file = file.path(project_root, 'data', 'baseline_feb10.RDS'))
baseline <- readRDS(
  file = file.path(project_root, 'data', 'baseline_feb10.RDS')
)

```

## Calibration

Target values are loaded in format long (`target.set`) and wide (`target.table.0`). Units are divided by 10,000 for convenience.

```{r calibration}
#### LOAD TARGET VALUES ####

target.list <- load.target(file.path(
  project_root,
  'data',
  'target_values.xlsx'
))
target.table <- target.list$wide
target.set <- target.list$long
target.vars <- target.set$var.label
target.set$target <- baseline$simulation[
  target.set$var,
  baseline$initial$pars['nPeriods', 'value']
]
target.set$ratio <- round(target.set$target / target.set$value, 2)

compute.target(baseline)
```

## Visualization

```{r variable_selection} 
#| fig-width: 7
#| fig-height: 4

# Variable Selection
selected.list <- list(
  Macroeconomic = c('c', 'va', 'go', 'id', 'gdef', 'cab', 'tb'),
  Social = c('n', 'nf', 'shp', 'shw'),
  Debt.and.Wealth = c('lh', 'lf', 'v', 'k', 'b_s'),
  Ecological = c('x_mat', 'land', 'mat', 'rec', 'emis', 'wa', 'water')
)


plot.selected.by_group(baseline, selected.list)
```

# Scenarios to Analyze

| Domain      | Shift              | From | To | Mechanism                  | Conserved Quantity            |
|-------------|--------------------|------|----|----------------------------|-------------------------------|
| Household Demand   | Green Diet Shift         | 7    | 8  | Final demand reallocation  | Total food expenditure        |
| Household Demand   | Green Energy Shift  | 31   | 32 | Final demand reallocation  | Energy service demand         |
| Production  | Wood circularity   | 11   | 12 | Technical coefficient change | Total wood material input   |
| Production  | Pulp circularity   | 13   | 14 | Technical coefficient change | Paper output                |
| Production  | Plastics circularity | 17 | 18 | Technical coefficient change | Plastic product output      |
| Production  | Metal circularity  | 26   | 27 | Technical coefficient change | Metal product output        |
| Production  | Glass circularity  | 21   | 22 | Technical coefficient change | Glass product output        |
| Production  | Cement circularity | 24   | 25 | Technical coefficient change | Cementitious output         |
| Production  | Green Energy Shift  | 31   | 32 | Technical coefficient change | Energy service to production|
| Production  | Construction reuse | 36   | 27 | Technical coefficient change | Construction output         |

# Scenario analysis (automated)

```{r scenario_loop}
#| results: asis
#| fig-width: 8
#| fig-height: 4

# ---- Scenario registry ----
scenarios_tbl <- tibble::tribble(
  ~shock, ~Domain,            ~Shift,                 ~From, ~To, ~Mechanism,                     ~Conserved_Quantity,
  1L,     "Household Demand", "Green Diet Shift",      7L,   8L, "Final demand reallocation",     "Total food expenditure",
  2L,     "Household Demand", "Green Energy Shift",   31L,  32L, "Final demand reallocation",     "Energy service demand",
  3L,     "Production",       "Wood circularity",     11L,  12L, "Technical coefficient change",  "Total wood material input",
  4L,     "Production",       "Pulp circularity",     13L,  14L, "Technical coefficient change",  "Paper output",
  5L,     "Production",       "Plastics circularity", 17L,  18L, "Technical coefficient change",  "Plastic product output",
  6L,     "Production",       "Metal circularity",    26L,  27L, "Technical coefficient change",  "Metal product output",
  7L,     "Production",       "Glass circularity",    21L,  22L, "Technical coefficient change",  "Glass product output",
  8L,     "Production",       "Cement circularity",   24L,  25L, "Technical coefficient change",  "Cementitious output",
  9L,     "Production",       "Green Energy Shift",   31L,  32L, "Technical coefficient change",  "Energy service to production",
  10L,    "Production",       "Construction reuse",   36L,  27L, "Technical coefficient change",  "Construction output"
) %>%
  mutate(
    display_name = paste0("Scenario ", shock, ": ", Domain, " - ", Shift),
    # Parameterize these as you implement new shocks.
    shock_vars = case_when(
      shock == 1L ~ "beta-7,beta-8",
      shock == 2L ~ "beta-31,beta-32",
      TRUE ~ NA_character_
    ),
    shock_sectors = case_when(
      shock == 1L ~ "7,8",
      shock == 2L ~ "31,32",
      TRUE ~ NA_character_
    ),
    # Where the scenario run is stored (RDS). You can adopt a consistent naming scheme.
    rds_file = case_when(
      shock == 1L ~ file.path(project_root, "data", "shock_1_diet_shock.RDS"),
      shock == 2L ~ file.path(project_root, "data", "shock_2_energy_shock.RDS"),
      TRUE ~ file.path(project_root, "data", paste0("shock_", shock, "_run.RDS"))
    ),
    # Where the scenario summary table is stored.
    table_file = case_when(
      shock == 1L ~ file.path(project_root, "data", "diet_shock_table.csv"),
      shock == 2L ~ file.path(project_root, "data", "energy_shock_table.csv"),
      TRUE ~ file.path(project_root, "data", paste0("shock_", shock, "_table.csv"))
    )
  )

# ---- Controls ----
run_if_missing <- FALSE  # set TRUE if you want quarto render to run missing scenarios via run.model()

# Time Period Selection (common to all scenarios)
t.shock <- initial$pars["t.shock", "value"]
t0 <- 55
tf <- initial$pars["nPeriods", "value"]

# Container to reuse later (e.g., comparison section)
scenario_runs <- list()
scenario_tables <- list()

plot_scenario <- function(s) {
  stopifnot(NROW(s) == 1L)

  shock <- s$shock
  shock.title <- s$display_name

  cat("

# ", shock.title, "

", sep = "")

  # Show the registry row so the doc is self-auditing
  cat(
    "- **Domain:** ", s$Domain, "
",
    "- **Mechanism:** ", s$Mechanism, "
",
    "- **From → To:** ", s$From, " → ", s$To, "
",
    "- **Conserved quantity:** ", s$Conserved_Quantity, "

",
    sep = ""
  )

  # Build initial shock
  initial_shock <- initial
  initial_shock$pars["shock", "value"] <- shock

  # Load (or optionally run) the scenario
  if (file.exists(s$rds_file)) {
    shock.run <- readRDS(s$rds_file)
  } else if (isTRUE(run_if_missing)) {
    shock.run <- run.model(
      initial_shock,
      mvp.model,
      log_file = file.path(project_root, "logs", paste0("shock_", shock, "_run.log")),
      log_append = FALSE,
      print_final_state = FALSE
    )
    saveRDS(shock.run, s$rds_file)
  } else {
    cat(
      "**Not run**: missing cached scenario file `", s$rds_file, "`.

",
      "Set `run_if_missing <- TRUE` in the chunk (and parameterize `shock_vars`/`shock_sectors`) to run it during render.

",
      sep = ""
    )
    return(invisible(NULL))
  }

  # ---- Visualize shocked variables (if specified) ----
  if (!is.na(s$shock_vars) && !is.na(s$shock_sectors)) {
    shock.vars <- strsplit(s$shock_vars, ",", fixed = TRUE)[[1]]
    shock.sectors <- as.integer(strsplit(s$shock_sectors, ",", fixed = TRUE)[[1]])

    cat("## Visualize shocked variables

")
    print(view.shock(shock.run, shock.vars = shock.vars, shock.sectors = shock.sectors))
  } else {
    cat("## Visualize shocked variables

")
    cat("*Pending parameterization for this scenario.*

")
  }

  # ---- Visualize impacts ----
  cat("## Visualize impacts

")
  df <- shock.long.new(baseline, shock.run)

  for (i in seq_along(selected.list)) {
    if ("gdef" %in% selected.list[[i]]) {
      print(view.vars(data = df, viz.vars = selected.list[[i]], var.label = names(selected.list)[i]))
    } else {
      print(view.scaled.vars(data = df, viz.vars = selected.list[[i]], var.label = names(selected.list)[i]))
    }
  }

  # ---- Selected variables plot ----
  cat("## Visualize selected variables

")
  selected <- c("n", "c", "shw", "va", "cab", "nf", "gdef", "tb", "go")

  p.selected.vars <- dplyr::filter(df, variable %in% selected) %>%
    ggplot(aes(x = time, y = value, color = area, linetype = scenario)) +
    facet_wrap(~name, scales = "free") +
    geom_line() +
    geom_vline(
      xintercept = initial$pars["t.shock", "value"],
      linetype = "dashed",
      linewidth = .4
    ) +
    labs(
      title = shock.title,
      subtitle = "Selected Indicators. Vertical dashed line indicates shock time"
    ) +
    theme(
      plot.title = element_text(face = "bold"),
      plot.subtitle = element_text(face = "italic", size = 10),
      legend.title = element_text(face = "bold")
    )

  print(p.selected.vars)

  # ---- Summary table + summary plots ----
  cat("## Summary visualization

")

  shock.table <- shock.summary(
    baseline,
    shock.run = shock.run,
    t_ = c(t.shock + 1, t.shock + 7, t.shock + 20),
    t_names = c("Immediate", "Short.Term", "Long.Term")
  )

  write.csv(shock.table, s$table_file, row.names = FALSE)

  # The helper script expects these filenames in the environment
  diet_shock_table_1_filename <- file.path(project_root, "figures", paste0("shock_", shock, "_table_1.pdf"))
  diet_shock_table_2_filename <- file.path(project_root, "figures", paste0("shock_", shock, "_table_2.pdf"))

  df <<- read.csv(s$table_file)
  source(file.path(project_root, "R", "scenario_analysis_2026.R"))

  # Expose the scenario objects for later use (comparison, etc.)
  scenario_runs[[as.character(shock)]] <<- shock.run
  scenario_tables[[as.character(shock)]] <<- shock.table

  # Backwards-compatibility for the existing comparison chunk
  if (shock == 1L) diet.table <<- shock.table
  if (shock == 2L) energy.table <<- shock.table

  invisible(NULL)
}

# Run only scenarios with existing cached RDS (or those you choose to run)
for (k in seq_len(nrow(scenarios_tbl))) {
  plot_scenario(scenarios_tbl[k, , drop = FALSE])
}

```


# Scenario Comparison

```{r comparison} 
#| fig-width: 10
#| fig-height: 21
#| fig-cap: "Comparison of Diet and Energy Transition Scenarios"

selected.list <- list(
  Macroeconomic = c('c', 'va', 'go', 'id', 'gdef', 'cab', 'tb'),
  Social = c('n', 'nf', 'shp', 'shw'),
  Debt.and.Wealth = c('lh', 'lf', 'v', 'k', 'b_s'),
  Ecological = c('x_mat', 'land', 'mat', 'rec', 'emis', 'wa', 'water')
)

lookup <- stack(selected.list) |>
  dplyr::rename(
    Variable = values,
    Name = ind
  )

scenario.tables <- list(diet = diet.table, energy = energy.table) %>%
  bind_rows(.id = "scenario") %>%
  mutate(
    scenario = factor(
      scenario,
      levels = c("diet", "energy"),
      labels = c('Household\nDiet Shift', 'Household\nEnergy Consumption Shift')
    )
  )

df <- scenario.tables |>
  dplyr::select(-Name) |> # drop old category column
  dplyr::left_join(lookup, by = "Variable") %>%
  filter(!is.na(Name))

source(file.path(project_root, "R", "scenario_comparison_2026.R"))
p.comparison
```

# Key Structural Findings from Scenario Comparison

The scenario analysis reveals a clear macro-ecological success combined with persistent social and distributional tensions. The green transition improves environmental indicators and external balances but generates labour market contraction and rising inequality, particularly over time.

## Macroeconomic Stabilization

Across both Core and Periphery, the green transition produces systematic macroeconomic improvements, especially beyond the immediate adjustment phase.

- Current account balance and trade balance improve in all time horizons  
- Government deficit narrows significantly in the short and long term  
- Value added and gross output recover after short-term adjustment costs  
- The Core captures stronger and earlier gains than the Periphery  

**Time dynamics:**  
Short-term disruption → Medium-term stabilization → Long-term consolidation  

These patterns are consistent with a structural reallocation process rather than a prolonged recessionary adjustment.

---

## Distributional Shift Toward Capital

The transition generates a persistent and intensifying redistribution from labour to capital.

| Indicator | Immediate | Short Term | Long Term | Interpretation |
|----------|-----------|------------|-----------|----------------|
| Wage share of income | ↓ | ↓↓ | ↓↓ | Labour loses income share |
| Profit share of income | ↑ | ↑↑ | ↑↑ | Capital captures transition rents |
| Worker disposable income | ↓ | ↓ | ↓ | Real labour income compression |
| Capitalist disposable income | ↑ | ↑↑ | ↑↑ | Income polarization widens |

The shift is not temporary; it strengthens over time, indicating a structural, capital-biased transition under existing institutional settings.

---

## Labour Market Contraction

Employment effects are consistently negative, with no clear recovery in the long term.

| Variable | Core | Periphery | Dynamics |
|----------|------|-----------|----------|
| Total employment | ↓ | slight ↓ | Persistent decline |
| Female employment | ↓↓ | slight ↓ | Gendered impact |
| Real consumption | ↓ | flat/slight ↓ | Demand compression |

The results suggest a labour-saving restructuring rather than a short-term displacement followed by reabsorption. The stronger fall in female employment points to sectoral reallocation away from feminized sectors.

---

## Core–Periphery Asymmetry

The macroeconomic benefits of the transition are unevenly distributed:

- The Core experiences stronger improvements in trade balances and current account positions  
- Financial variables and wealth indicators suggest capital re-accumulation in the Core  
- The Periphery benefits less from value-added gains  

This pattern indicates upstream green upgrading concentrated in core economies, with limited peripheral value capture.

---

## Strong and Immediate Ecological Gains

Environmental improvements are large, immediate, and sustained across all horizons.

| Indicator | Pattern |
|----------|---------|
| CO₂ emissions | Large decline |
| Material extraction | Decline |
| Land use | Decline |
| Waste | Decline |
| Water use | Decline |
| Recycling | Increase |

Unlike the social variables, ecological gains are front-loaded rather than delayed.

---

## Time Horizon Summary

| Dimension | Immediate | Short Term | Long Term |
|-----------|-----------|------------|-----------|
| Ecology | Large gains | Maintained | Maintained |
| Macroeconomic stability | Mixed | Improves | Consolidates |
| Employment | Falls | Falls further | No recovery |
| Inequality | Rises | Rises more | Structural shift |
| Core–Periphery gap | Limited | Widening | Entrenched |

---

## Overall Interpretation

The modeled green transition is:

**Ecologically effective**  
**Macroeconomically stabilizing**  
**Socially regressive under current institutions**  

Without complementary policies such as:

- Wage protection and collective bargaining support  
- Employment guarantees or green job programs  
- Labour-absorbing industrial policy  
- Gender-sensitive transition measures  

the transition path risks becoming politically destabilizing despite ecological success.

---

## Executive Takeaway

- The green transition delivers rapid ecological improvements and medium-term macroeconomic stabilization, particularly in external balances and fiscal positions. 
- However, these gains are accompanied by persistent labour displacement, declining wage shares, rising profit shares, and falling worker incomes, with especially strong negative effects on female employment. 
- Inequality widens over time, and the Core captures a disproportionate share of the macroeconomic benefits relative to the Periphery.
- Without strong redistributive and labour-market institutions, the transition appears environmentally effective but socially and politically fragile.

