---
title: "Scenario Analysis"
subtitle: "Insights from a Core-Periphery SOCIO-ECO-MRIO-SFC Model"
author:
  - name: "Oriol Vallès Codina"
    affiliation: "Johns Hopkins University, Net Zero Industrial Policy Lab, Department of Political Science"
    email: 
      - "oriolvallescodina@gmail.com"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc_float: true
    self-contained: true
    page-layout: full
    css: fullwidth.css
execute:
  echo: false
  working-dir: ..
---

# Initialization

The initial values and variables are loaded from the file stored in variable `identif` in a different structure as before.

```{r packages, warning = F, message = F}
library(tidyverse)
library(igraph)
library(qgraph)
library(visNetwork)
library(scales)
```

```{r init, warning = F}
# -----------------------------
# Robust project root detection
# -----------------------------
qmd_in <- knitr::current_input()

if (!is.null(qmd_in) && nzchar(qmd_in)) {
  qmd_path <- normalizePath(qmd_in, winslash = "/")
  root <- normalizePath(file.path(dirname(qmd_path), ".."), winslash = "/")
  root_method <- "knitr::current_input()"
} else {
  # Fallback for interactive execution / sourcing
  wd0 <- normalizePath(getwd(), winslash = "/")
  root <- if (basename(wd0) == "qmd") {
    normalizePath("..", winslash = "/")
  } else {
    wd0
  }
  root_method <- "getwd() fallback"
}

dir_data <- file.path(root, "data")

cat("root_method:", root_method, "\n")
cat("getwd():", normalizePath(getwd(), winslash = "/"), "\n")
cat("root  :", root, "\n")
cat("dir_data:", dir_data, "\n")
cat(
  "VarDef exists?:",
  file.exists(file.path(dir_data, "Variable_Definitions.csv")),
  "\n"
)
cat(
  "initial exists?:",
  file.exists(file.path(dir_data, "initial_state_2026.xlsx")),
  "\n"
)

# -----------------------------
# Canonical directories
# -----------------------------
model_dir <- file.path(root, "model")
functions_dir <- file.path(root, "functions") # keep only if you still use it
utils_dir <- file.path(root, "utils")

# -----------------------------
# Core inputs
# -----------------------------
variable.table <- read.csv(
  file.path(dir_data, "Variable_Definitions.csv"),
  stringsAsFactors = FALSE
)

initial_filename <- file.path(dir_data, "initial_state_2026.xlsx")
stopifnot(file.exists(initial_filename))

sector_list <- read.csv(
  file.path(dir_data, "sector_list.csv"),
  stringsAsFactors = FALSE
) %>%
  mutate(label = gsub("<br>", " ", label))

# -----------------------------
# Scenario workspace dirs
# -----------------------------
workspace_dir <- file.path(root, "output", "scenarios")
dir.create(workspace_dir, showWarnings = FALSE, recursive = TRUE)

dir_runs <- file.path(workspace_dir, "shock_runs")
dir_tables <- file.path(workspace_dir, "shock_tables")
dir_logs <- file.path(workspace_dir, "logs")
dir_long <- file.path(workspace_dir, "shock_long")
dir_cache <- file.path(workspace_dir, "scenario_cache")
dir_rds <- file.path(workspace_dir, "rds_files")

invisible(lapply(
  list(dir_runs, dir_tables, dir_logs, dir_long, dir_cache, dir_rds),
  dir.create,
  showWarnings = FALSE,
  recursive = TRUE
))

# -----------------------------
# Source utilities + model code
# -----------------------------
source(file.path(utils_dir, "utils.R"), local = FALSE)

# Load initial state
initial <- load.init(initial_filename)

# Model functions
source(file.path(model_dir, "run_model_2026.R"), local = FALSE)
source(file.path(model_dir, "MVP_model_2026.R"), local = FALSE)

# Scenario analysis helpers
source(file.path(utils_dir, "scenario_analysis_2026.R"), local = FALSE)
```

```{r filenames}
# ---- Workspace filenames (generated / cached) ----
baseline_filename <- file.path(workspace_dir, "baseline_2026.RDS")
target_filename <- file.path(workspace_dir, "target_values.xlsx")
scenario_filename <- file.path(workspace_dir, "scenarios.csv")

get_shock_filename <- function(n) {
  file.path(dir_runs, paste0("shock_", n, "_run.RDS"))
}
get_table_filename <- function(n) {
  file.path(dir_tables, paste0("shock_", n, "_table.csv"))
}
get_log_filename <- function(n) {
  file.path(dir_logs, paste0("shock_", n, "_run.log"))
}
```

```{r theme}
theme_leeds <- function() {
  theme_minimal() +
    theme(
      plot.title = element_text(face = "bold", size = 14),
      legend.title = element_text(face = "bold"),
      strip.text.y = element_text(angle = 0),
      plot.subtitle = element_text(face = "italic", size = 10),
      strip.text.y.right = element_text(angle = 0)
    )
}

theme_set(
  theme_leeds()
)
```


# Baseline 

```{r simulation, warning = F, message = F}
# baseline <- run.model(
#   initial,
#   mvp.model,
#   log_file = file.path(root, "logs", "baseline_run.log"),
#   log_append = FALSE,
#   print_final_state = FALSE
# )

# saveRDS(baseline, file = file.path(root, 'data', 'baseline_feb10.RDS'))
baseline <- readRDS(
  file = baseline_filename
)

```

## Calibration

Target values are loaded in format long (`target.set`) and wide (`target.table.0`). Units are divided by 10,000 for convenience.

```{r calibration}
#### LOAD TARGET VALUES ####

target.list <- load.target(target_filename, variable.table)
target.table <- target.list$wide
target.set <- target.list$long
target.vars <- target.set$var.label
target.set$target <- baseline$simulation[
  target.set$var,
  baseline$initial$pars['nPeriods', 'value']
]
target.set$ratio <- round(target.set$target / target.set$value, 2)

compute.target(baseline)
```

## Visualization

```{r variable_selection} 
#| fig-width: 7
#| fig-height: 4

# Variable Selection
selected.list <- list(
  Macroeconomic = c('c', 'va', 'go', 'id', 'gdef', 'cab', 'tb'),
  Social = c('n', 'nf', 'shp', 'shw'),
  Debt.and.Wealth = c('lh', 'lf', 'v', 'k', 'b_s'),
  Ecological = c('x_mat', 'land', 'mat', 'rec', 'emis', 'wa', 'water')
)

plot.selected.by_group(baseline, selected.list)
```

___

# Scenarios to Analyze

```{r scenarios.table}
sc <- read.csv(file = scenario_filename)

sc %>%
  select(-display_name) %>%
  kable(
    caption = "Scenario Registry",
    align = "l",
    escape = TRUE
  ) %>%
  kable_styling(
    full_width = FALSE,
    bootstrap_options = c("striped", "hover", "condensed"),
    font_size = 12
  ) %>%
  scroll_box(width = "100%", height = "600px")

scenario_list <- list()

# Time Period Selection
t.shock <- initial$pars['t.shock', 'value'] # shock time period
t0 <- 55 # initial time period for visualization
tf <- initial$pars['nPeriods', 'value'] # final time period for visualization
```

___

# Scenario 1: Household Demand - Green Diet Shift

```{r shock.run}
# Shock Selection and Run
initial_shock <- initial
n_shock <- initial_shock$pars['shock', 'value'] <- 1

scenario_list[[n_shock]] <- scenario <- run_or_load_shock(
  n_shock,
  initial_shock,
  mvp.model,
  force = FALSE
)
```

## Visualize Shocked Variables

```{r shock.view}
#| fig-width: 6
#| fig-height: 2

view.shock(
  scenario,
  shock.vars = c(
    'beta-7',
    'beta-8'
  ),
  shock.sectors = c(7, 8),
  shock.title = sc$display_name[n_shock]
)
```

## Visualize Impacts

```{r shock.visualize.scaled}
#| fig-width: 8
#| fig-height: 4

p.selected.vars <- plot_selected_vars(
  baseline = baseline,
  scenario = scenario,
  selected.list = selected.list,
  shock_title = sc$display_name[n_shock],
  t_shock = initial$pars["t.shock", "value"],
  print_blocks = TRUE
)

p.selected.vars
```

```{r summary.table.1}
#| fig-width: 10
#| fig-height: 10

scenario.table <- shock.summary(
  baseline,
  shock.run = scenario,
  t_ = c(t.shock + 1, t.shock + 7, t.shock + 20),
  t_names = c('Immediate', 'Short.Term', 'Long.Term')
)

# kable(diet.table, booktabs = T)
write.csv(scenario.table, get_table_filename(n_shock), row.names = FALSE)

scenario.table <- read.csv(get_table_filename(n_shock))

scenario_analysis_2026_plot(
  shock_df = scenario.table,
  shock_title = sc$display_name[n_shock],
  variable.table = variable.table
)
```

___

# Scenario 2: Household Demand - Green Energy Shift

```{r shock.run.2}
# Shock Selection and Run
initial_shock <- initial
n_shock <- initial_shock$pars['shock', 'value'] <- 2

scenario_list[[n_shock]] <- scenario <- run_or_load_shock(
  n_shock,
  initial_shock,
  mvp.model,
  force = FALSE
)
```

## Visualize Shocked Variables

```{r shock.view.2}
#| fig-width: 6
#| fig-height: 2

view.shock(
  scenario,
  shock.vars = c(
    'beta-31',
    'beta-32'
  ),
  shock.sectors = c(31, 32),
  shock.title = sc$display_name[n_shock]
)
```

## Visualize Impacts

```{r shock.visualize.scaled.2}
#| fig-width: 8
#| fig-height: 4

p.selected.vars <- plot_selected_vars(
  baseline = baseline,
  scenario = scenario,
  selected.list = selected.list,
  shock_title = sc$display_name[n_shock],
  t_shock = initial$pars["t.shock", "value"],
  print_blocks = TRUE
)

p.selected.vars
```

```{r summary.table.2}
#| fig-width: 10
#| fig-height: 10

scenario.table <- shock.summary(
  baseline,
  shock.run = scenario,
  t_ = c(t.shock + 1, t.shock + 7, t.shock + 20),
  t_names = c('Immediate', 'Short.Term', 'Long.Term')
)

# kable(energy.table, booktabs = T)
write.csv(scenario.table, get_table_filename(n_shock), row.names = FALSE)

scenario.table <- read.csv(get_table_filename(n_shock))

scenario_analysis_2026_plot(
  shock_df = scenario.table,
  shock_title = sc$display_name[n_shock],
  variable.table = variable.table
)
```

___

# Scenario 7: Production - Wood Shift

```{r shock.run.7}
# Shock Selection and Run
initial_shock <- initial
n_shock <- initial_shock$pars['shock', 'value'] <- 7

scenario_list[[n_shock]] <- scenario <- run_or_load_shock(
  n_shock,
  initial_shock,
  mvp.model,
  force = FALSE
)
```

## Visualize Shocked Coefficients

```{r shock.view.7}
#| fig-width: 6
#| fig-height: 4

pairs <- data.frame(
  row = sc %>%
    filter(shock == n_shock) %>%
    select(from, to) %>%
    rep(2) %>%
    unlist,
  col = c(31, 31, 32, 32)
)

view.A(
  scenario,
  pairs,
  title = paste(
    sc %>% filter(shock == n_shock) %>% pull(shift),
    "input-output coefficients"
  )
)
```

## Visualize Impacts

```{r shock.visualize.scaled.7}
#| fig-width: 8
#| fig-height: 4

p.selected.vars <- plot_selected_vars(
  baseline = baseline,
  scenario = scenario,
  selected.list = selected.list,
  shock_title = sc$display_name[n_shock],
  t_shock = initial$pars["t.shock", "value"],
  print_blocks = TRUE
)

p.selected.vars
```

## Summary Visualization

```{r summary.table.7}
#| fig-width: 10
#| fig-height: 10

scenario.table <- shock.summary(
  baseline,
  shock.run = scenario,
  t_ = c(t.shock + 1, t.shock + 7, t.shock + 20),
  t_names = c('Immediate', 'Short.Term', 'Long.Term')
)

# kable(shock_3_table, booktabs = T)
write.csv(scenario.table, get_table_filename(n_shock), row.names = FALSE)

scenario.table <- read.csv(get_table_filename(n_shock))

scenario_analysis_2026_plot(
  shock_df = scenario.table,
  shock_title = sc$display_name[n_shock],
  variable.table = variable.table
)

```

___

# Scenario 8: Production - Pulp Shift

```{r shock.run.8}
# Shock Selection and Run
initial_shock <- initial
n_shock <- initial_shock$pars['shock', 'value'] <- 8

scenario_list[[n_shock]] <- scenario <- run_or_load_shock(
  n_shock,
  initial_shock,
  mvp.model,
  force = FALSE
)
```

## Visualize Shocked Coefficients

```{r shock.view.8}
#| fig-width: 6
#| fig-height: 4

pairs <- data.frame(
  row = sc %>%
    filter(shock == n_shock) %>%
    select(from, to) %>%
    rep(2) %>%
    unlist,
  col = c(31, 31, 32, 32)
)

view.A(
  scenario,
  pairs,
  title = paste(
    sc %>% filter(shock == n_shock) %>% pull(shift),
    "input-output coefficients"
  )
)
```

## Visualize Impacts

```{r shock.visualize.scaled.8}
#| fig-width: 8
#| fig-height: 4

p.selected.vars <- plot_selected_vars(
  baseline = baseline,
  scenario = scenario,
  selected.list = selected.list,
  shock_title = sc$display_name[n_shock],
  t_shock = initial$pars["t.shock", "value"],
  print_blocks = TRUE
)

p.selected.vars
```

## Summary Visualization

```{r summary.table.8}
#| fig-width: 10
#| fig-height: 10

scenario.table <- shock.summary(
  baseline,
  shock.run = scenario,
  t_ = c(t.shock + 1, t.shock + 7, t.shock + 20),
  t_names = c('Immediate', 'Short.Term', 'Long.Term')
)

write.csv(scenario.table, get_table_filename(n_shock), row.names = FALSE)

scenario.table <- read.csv(get_table_filename(n_shock))

scenario_analysis_2026_plot(
  shock_df = scenario.table,
  shock_title = sc$display_name[n_shock],
  variable.table = variable.table
)

```

___

# Scenario 9: Production - Plastics Shift

```{r shock.run.9}
# Shock Selection and Run
initial_shock <- initial
n_shock <- initial_shock$pars['shock', 'value'] <- 9

scenario_list[[n_shock]] <- scenario <- run_or_load_shock(
  n_shock,
  initial_shock,
  mvp.model,
  force = FALSE
)
```

## Visualize Shocked Coefficients

```{r shock.view.9}
#| fig-width: 6
#| fig-height: 4

pairs <- data.frame(
  row = sc %>%
    filter(shock == n_shock) %>%
    select(from, to) %>%
    rep(2) %>%
    unlist,
  col = c(31, 31, 32, 32)
)

view.A(
  scenario,
  pairs,
  title = paste(
    sc %>% filter(shock == n_shock) %>% pull(shift),
    "input-output coefficients"
  )
)
```

## Visualize Impacts

```{r shock.visualize.scaled.9}
#| fig-width: 8
#| fig-height: 4

p.selected.vars <- plot_selected_vars(
  baseline = baseline,
  scenario = scenario,
  selected.list = selected.list,
  shock_title = sc$display_name[n_shock],
  t_shock = initial$pars["t.shock", "value"],
  print_blocks = TRUE
)

p.selected.vars
```

## Summary Visualization

```{r summary.table.9}
#| fig-width: 10
#| fig-height: 10

scenario.table <- shock.summary(
  baseline,
  shock.run = scenario,
  t_ = c(t.shock + 1, t.shock + 7, t.shock + 20),
  t_names = c('Immediate', 'Short.Term', 'Long.Term')
)

write.csv(scenario.table, get_table_filename(n_shock), row.names = FALSE)

scenario.table <- read.csv(get_table_filename(n_shock))

scenario_analysis_2026_plot(
  shock_df = scenario.table,
  shock_title = sc$display_name[n_shock],
  variable.table = variable.table
)

```

___ 

# Scenario 10: Production - Metal Shift

```{r shock.run.10}
# Shock Selection and Run
initial_shock <- initial
n_shock <- initial_shock$pars['shock', 'value'] <- 10

scenario_list[[n_shock]] <- scenario <- run_or_load_shock(
  n_shock,
  initial_shock,
  mvp.model,
  force = FALSE
)
```

## Visualize Shocked Coefficients

```{r shock.view.10}
#| fig-width: 6
#| fig-height: 4

pairs <- data.frame(
  row = sc %>%
    filter(shock == n_shock) %>%
    select(from, to) %>%
    rep(2) %>%
    unlist,
  col = c(31, 31, 32, 32)
)

view.A(
  scenario,
  pairs,
  title = paste(
    sc %>% filter(shock == n_shock) %>% pull(shift),
    "input-output coefficients"
  )
)
```

## Visualize Impacts

```{r shock.visualize.scaled.10}
#| fig-width: 8
#| fig-height: 4

p.selected.vars <- plot_selected_vars(
  baseline = baseline,
  scenario = scenario,
  selected.list = selected.list,
  shock_title = sc$display_name[n_shock],
  t_shock = initial$pars["t.shock", "value"],
  print_blocks = TRUE
)

p.selected.vars
```

## Summary Visualization

```{r summary.table.10}
#| fig-width: 10
#| fig-height: 10

scenario.table <- shock.summary(
  baseline,
  shock.run = scenario,
  t_ = c(t.shock + 1, t.shock + 7, t.shock + 20),
  t_names = c('Immediate', 'Short.Term', 'Long.Term')
)

write.csv(scenario.table, get_table_filename(n_shock), row.names = FALSE)

scenario.table <- read.csv(get_table_filename(n_shock))

scenario_analysis_2026_plot(
  shock_df = scenario.table,
  shock_title = sc$display_name[n_shock],
  variable.table = variable.table
)

```

___

# Scenario 11: Production - Glass Shift

```{r shock.run.11}
# Shock Selection and Run
initial_shock <- initial
n_shock <- initial_shock$pars['shock', 'value'] <- 11

scenario_list[[n_shock]] <- scenario <- run_or_load_shock(
  n_shock,
  initial_shock,
  mvp.model,
  force = FALSE
)
```

## Visualize Shocked Coefficients

```{r shock.view.11}
#| fig-width: 6
#| fig-height: 4

pairs <- data.frame(
  row = sc %>%
    filter(shock == n_shock) %>%
    select(from, to) %>%
    rep(2) %>%
    unlist,
  col = c(31, 31, 32, 32)
)

view.A(
  scenario,
  pairs,
  title = paste(
    sc %>% filter(shock == n_shock) %>% pull(shift),
    "input-output coefficients"
  )
)
```

## Visualize Impacts

```{r shock.visualize.scaled.11}
#| fig-width: 8
#| fig-height: 4

p.selected.vars <- plot_selected_vars(
  baseline = baseline,
  scenario = scenario,
  selected.list = selected.list,
  shock_title = sc$display_name[n_shock],
  t_shock = initial$pars["t.shock", "value"],
  print_blocks = TRUE
)

p.selected.vars
```

## Summary Visualization

```{r summary.table.11}
#| fig-width: 10
#| fig-height: 10

scenario.table <- shock.summary(
  baseline,
  shock.run = scenario,
  t_ = c(t.shock + 1, t.shock + 7, t.shock + 20),
  t_names = c('Immediate', 'Short.Term', 'Long.Term')
)

write.csv(scenario.table, get_table_filename(n_shock), row.names = FALSE)

scenario.table <- read.csv(get_table_filename(n_shock))

scenario_analysis_2026_plot(
  shock_df = scenario.table,
  shock_title = sc$display_name[n_shock],
  variable.table = variable.table
)

```

___

# Scenario 12: Production - Cement Shift

```{r shock.run.12}
# Shock Selection and Run
initial_shock <- initial
n_shock <- initial_shock$pars['shock', 'value'] <- 11

scenario_list[[n_shock]] <- scenario <- run_or_load_shock(
  n_shock,
  initial_shock,
  mvp.model,
  force = FALSE
)
```

## Visualize Shocked Coefficients

```{r shock.view.12}
#| fig-width: 6
#| fig-height: 4

pairs <- data.frame(
  row = sc %>%
    filter(shock == n_shock) %>%
    select(from, to) %>%
    rep(2) %>%
    unlist,
  col = c(31, 31, 32, 32)
)

view.A(
  scenario,
  pairs,
  title = paste(
    sc %>% filter(shock == n_shock) %>% pull(shift),
    "input-output coefficients"
  )
)
```

## Visualize Impacts

```{r shock.visualize.scaled.12}
#| fig-width: 8
#| fig-height: 4

p.selected.vars <- plot_selected_vars(
  baseline = baseline,
  scenario = scenario,
  selected.list = selected.list,
  shock_title = sc$display_name[n_shock],
  t_shock = initial$pars["t.shock", "value"],
  print_blocks = TRUE
)

p.selected.vars
```

## Summary Visualization


```{r summary.table.12}
#| fig-width: 10
#| fig-height: 10

scenario.table <- shock.summary(
  baseline,
  shock.run = scenario,
  t_ = c(t.shock + 1, t.shock + 7, t.shock + 20),
  t_names = c('Immediate', 'Short.Term', 'Long.Term')
)

write.csv(scenario.table, get_table_filename(n_shock), row.names = FALSE)

scenario.table <- read.csv(get_table_filename(n_shock))

scenario_analysis_2026_plot(
  shock_df = scenario.table,
  shock_title = sc$display_name[n_shock],
  variable.table = variable.table
)

```


___

# Scenario 13: Production - Energy Shift

```{r shock.run.13}
# Shock Selection and Run
initial_shock <- initial
n_shock <- initial_shock$pars['shock', 'value'] <- 13

scenario_list[[n_shock]] <- scenario <- run_or_load_shock(
  n_shock,
  initial_shock,
  mvp.model,
  force = FALSE
)
```

## Visualize Shocked Coefficients

```{r shock.view.13}
#| fig-width: 6
#| fig-height: 4

pairs <- data.frame(
  row = sc %>%
    filter(shock == n_shock) %>%
    select(from, to) %>%
    rep(2) %>%
    unlist,
  col = c(31, 31, 32, 32)
)

view.A(
  scenario,
  pairs,
  title = paste(
    sc %>% filter(shock == n_shock) %>% pull(shift),
    "input-output coefficients"
  )
)
```

## Visualize Impacts

```{r shock.visualize.scaled.13}
#| fig-width: 8
#| fig-height: 4

p.selected.vars <- plot_selected_vars(
  baseline = baseline,
  scenario = scenario,
  selected.list = selected.list,
  shock_title = sc$display_name[n_shock],
  t_shock = initial$pars["t.shock", "value"],
  print_blocks = TRUE
)

p.selected.vars
```

## Summary Visualization

```{r summary.table.13}
#| fig-width: 10
#| fig-height: 10

scenario.table <- shock.summary(
  baseline,
  shock.run = scenario,
  t_ = c(t.shock + 1, t.shock + 7, t.shock + 20),
  t_names = c('Immediate', 'Short.Term', 'Long.Term')
)

write.csv(scenario.table, get_table_filename(n_shock), row.names = FALSE)

scenario.table <- read.csv(get_table_filename(n_shock))

scenario_analysis_2026_plot(
  shock_df = scenario.table,
  shock_title = sc$display_name[n_shock],
  variable.table = variable.table
)

```


___

# Scenario 14: Production - Construction Shift

```{r shock.run.14}
# Shock Selection and Run
initial_shock <- initial
n_shock <- initial_shock$pars['shock', 'value'] <- 14

scenario_list[[n_shock]] <- scenario <- run_or_load_shock(
  n_shock,
  initial_shock,
  mvp.model,
  force = FALSE
)
```

## Visualize Shocked Coefficients

```{r shock.view.14}
#| fig-width: 6
#| fig-height: 4

pairs <- data.frame(
  row = sc %>%
    filter(shock == n_shock) %>%
    select(from, to) %>%
    rep(2) %>%
    unlist,
  col = c(31, 31, 32, 32)
)

view.A(
  scenario,
  pairs,
  title = paste(
    sc %>% filter(shock == n_shock) %>% pull(shift),
    "input-output coefficients"
  )
)
```

## Visualize Impacts

```{r shock.visualize.scaled.14}
#| fig-width: 8
#| fig-height: 4

p.selected.vars <- plot_selected_vars(
  baseline = baseline,
  scenario = scenario,
  selected.list = selected.list,
  shock_title = sc$display_name[n_shock],
  t_shock = initial$pars["t.shock", "value"],
  print_blocks = TRUE
)

p.selected.vars
```

## Summary Visualization

```{r summary.table.14}
#| fig-width: 10
#| fig-height: 10

scenario.table <- shock.summary(
  baseline,
  shock.run = scenario,
  t_ = c(t.shock + 1, t.shock + 7, t.shock + 20),
  t_names = c('Immediate', 'Short.Term', 'Long.Term')
)

write.csv(scenario.table, get_table_filename(n_shock), row.names = FALSE)

scenario.table <- read.csv(get_table_filename(n_shock))

scenario_analysis_2026_plot(
  shock_df = scenario.table,
  shock_title = sc$display_name[n_shock],
  variable.table = variable.table
)

```

___

# Scenario Comparison

```{r comparison} 
#| fig-width: 10
#| fig-height: 21
#| fig-cap: "Comparison of Diet and Energy Transition Scenarios"

selected.list <- list(
  Macroeconomic = c('c', 'va', 'go', 'id', 'gdef', 'cab', 'tb'),
  Social = c('n', 'nf', 'shp', 'shw'),
  Debt.and.Wealth = c('lh', 'lf', 'v', 'k', 'b_s'),
  Ecological = c('x_mat', 'land', 'mat', 'rec', 'emis', 'wa', 'water')
)

lookup <- stack(selected.list) |>
  dplyr::rename(
    Variable = values,
    Name = ind
  )

# 1) Give scenario_list stable names (use display_name)
names(scenario_list) <- sc$display_name[1:length(scenario_list)]

# 2) Build tables
scenario_tables_list <- lapply(scenario_list, function(x) {
  shock.summary(
    baseline,
    shock.run = x,
    t_ = c(t.shock + 1, t.shock + 7, t.shock + 20),
    t_names = c("Immediate", "Short.Term", "Long.Term")
  )
})

# 3) Keep the SAME names (display_name). Do NOT rename to shift.
# names(scenario_tables_list) <- ...  # <- delete this line

# 4) Bind into one data frame; factor levels follow the list order
scenario.tables <- scenario_tables_list %>%
  bind_rows(.id = "scenario") %>%
  mutate(
    scenario = factor(scenario, levels = names(scenario_tables_list))
  )

df <- scenario.tables |>
  dplyr::select(-Name) |> # drop old category column
  dplyr::left_join(lookup, by = "Variable") %>%
  filter(!is.na(Name))


source(file.path(utils_dir, 'scenario_comparison_2026.R'))

p.comparison
```

___

# Key Structural Findings from Scenario Comparison

The scenario analysis reveals a clear macro-ecological success combined with persistent social and distributional tensions. The green transition improves environmental indicators and external balances but generates labour market contraction and rising inequality, particularly over time.

## Macroeconomic Stabilization

Across both Core and Periphery, the green transition produces systematic macroeconomic improvements, especially beyond the immediate adjustment phase.

- Current account balance and trade balance improve in all time horizons  
- Government deficit narrows significantly in the short and long term  
- Value added and gross output recover after short-term adjustment costs  
- The Core captures stronger and earlier gains than the Periphery  

**Time dynamics:**  
Short-term disruption → Medium-term stabilization → Long-term consolidation  

These patterns are consistent with a structural reallocation process rather than a prolonged recessionary adjustment.

---

## Distributional Shift Toward Capital

The transition generates a persistent and intensifying redistribution from labour to capital.

| Indicator | Immediate | Short Term | Long Term | Interpretation |
|----------|-----------|------------|-----------|----------------|
| Wage share of income | ↓ | ↓↓ | ↓↓ | Labour loses income share |
| Profit share of income | ↑ | ↑↑ | ↑↑ | Capital captures transition rents |
| Worker disposable income | ↓ | ↓ | ↓ | Real labour income compression |
| Capitalist disposable income | ↑ | ↑↑ | ↑↑ | Income polarization widens |

The shift is not temporary; it strengthens over time, indicating a structural, capital-biased transition under existing institutional settings.

---

## Labour Market Contraction

Employment effects are consistently negative, with no clear recovery in the long term.

| Variable | Core | Periphery | Dynamics |
|----------|------|-----------|----------|
| Total employment | ↓ | slight ↓ | Persistent decline |
| Female employment | ↓↓ | slight ↓ | Gendered impact |
| Real consumption | ↓ | flat/slight ↓ | Demand compression |

The results suggest a labour-saving restructuring rather than a short-term displacement followed by reabsorption. The stronger fall in female employment points to sectoral reallocation away from feminized sectors.

---

## Core–Periphery Asymmetry

The macroeconomic benefits of the transition are unevenly distributed:

- The Core experiences stronger improvements in trade balances and current account positions  
- Financial variables and wealth indicators suggest capital re-accumulation in the Core  
- The Periphery benefits less from value-added gains  

This pattern indicates upstream green upgrading concentrated in core economies, with limited peripheral value capture.

---

## Strong and Immediate Ecological Gains

Environmental improvements are large, immediate, and sustained across all horizons.

| Indicator | Pattern |
|----------|---------|
| CO₂ emissions | Large decline |
| Material extraction | Decline |
| Land use | Decline |
| Waste | Decline |
| Water use | Decline |
| Recycling | Increase |

Unlike the social variables, ecological gains are front-loaded rather than delayed.

---

## Time Horizon Summary

| Dimension | Immediate | Short Term | Long Term |
|-----------|-----------|------------|-----------|
| Ecology | Large gains | Maintained | Maintained |
| Macroeconomic stability | Mixed | Improves | Consolidates |
| Employment | Falls | Falls further | No recovery |
| Inequality | Rises | Rises more | Structural shift |
| Core–Periphery gap | Limited | Widening | Entrenched |

---

## Overall Interpretation

The modeled green transition is:

**Ecologically effective**  
**Macroeconomically stabilizing**  
**Socially regressive under current institutions**  

Without complementary policies such as:

- Wage protection and collective bargaining support  
- Employment guarantees or green job programs  
- Labour-absorbing industrial policy  
- Gender-sensitive transition measures  

the transition path risks becoming politically destabilizing despite ecological success.

---

## Executive Takeaway

- The green transition delivers rapid ecological improvements and medium-term macroeconomic stabilization, particularly in external balances and fiscal positions. 
- However, these gains are accompanied by persistent labour displacement, declining wage shares, rising profit shares, and falling worker incomes, with especially strong negative effects on female employment. 
- Inequality widens over time, and the Core captures a disproportionate share of the macroeconomic benefits relative to the Periphery.
- Without strong redistributive and labour-market institutions, the transition appears environmentally effective but socially and politically fragile.

