---
title: "Scenario Key Findings"
subtitle: "EU–RoW MRIO-SFC Model | Circular Economy Transition Scenarios"
author:
  - name: "Oriol Vallès Codina"
    affiliation: "Johns Hopkins University, Net Zero Industrial Policy Lab, Department of Political Science"
    email:
      - "oriolvallescodina@gmail.com"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-float: true
    self-contained: true
    page-layout: full
    css: fullwidth.css
  docx:
    toc: true
prefer-html: true
execute:
  echo: false
  working-dir: ..
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
```

```{r load-rds, include=FALSE}
qmd_in <- knitr::current_input()
if (!is.null(qmd_in) && nzchar(qmd_in)) {
  qmd_path <- normalizePath(qmd_in, winslash = "/")
  root <- normalizePath(file.path(dirname(qmd_path), ".."), winslash = "/")
} else {
  wd0  <- normalizePath(getwd(), winslash = "/")
  root <- if (basename(wd0) == "qmd") normalizePath("..", winslash = "/") else wd0
}

sc <- read.csv(file.path(root, "data", "scenarios.csv"))

baseline_path <- file.path(root, "output", "scenarios", "full_mrio_baseline_2026.RDS")
if (!file.exists(baseline_path))
  baseline_path <- file.path(root, "output", "scenarios", "baseline_2026.RDS")
base <- readRDS(baseline_path)

t_lt <- 90L
t_st <- 77L
t_im <- 71L

pct_dev <- function(run, var, t) {
  (run$simulation[var, t] / base$simulation[var, t] - 1) * 100
}

pct_vec <- function(run, vars, t) {
  sapply(vars, function(v) if (is.null(run)) NA_real_ else pct_dev(run, v, t))
}

loads <- lapply(seq_len(nrow(sc)), function(n) {
  f <- file.path(root, "output", "scenarios", "shock_runs",
                 paste0("shock_", n, "_run.RDS"))
  if (!file.exists(f)) return(NULL)
  readRDS(f)
})
names(loads) <- sc$display_name
```

---

# Scenarios

```{r scenario-table}
sc |>
  dplyr::select(shock, display_name, domain, sector, shift, rho) |>
  dplyr::rename(
    `#` = shock,
    Scenario = display_name,
    Domain = domain,
    Channel = sector,
    Material = shift,
    `ρ` = rho
  ) |>
  knitr::kable(caption = "Scenario Registry", align = "l", booktabs = TRUE)
```

---

# Scenario Comparison

The comparison figures (generated by `R/export_comparison_plots.R`) present all 14 scenarios simultaneously across three time horizons: Immediate (*t*+1), Short-Term (*t*+7), and Long-Term (*t*+20). The integrated figure (08/09) organises variables into five categories across two columns: Macroeconomic, Social, and Ecological on the left; Financial and Circular Economy sector outputs on the right. Separate figures are produced for EU (domestic) and RoW (cross-border) effects.

---

# Summary of Macroeconomic, Social, and Ecological Implications

## The Portfolio Scale Principle

The most robust finding across all 14 scenarios is that the rank ordering of outcomes is predicted by the absolute scale of the targeted demand channel, not by material-specific multiplier variation or sectoral intensity alone. The structural Leontief factor (l₁₂ − l₁₁) is approximately −1 across all material sectors, meaning that primary-to-secondary substitution consistently contracts aggregate demand in proportion to channel size. What varies decisively across scenarios is the size of the demand portfolio the policy instrument engages.

This generates a clear ranking:

- **Food (Sc1)** produces the largest final-demand employment contraction (Z1 *n* = −0.09% LT) because household food demand (23.7 B EUR) is the largest household material channel. It also drives the largest primary material extraction reduction among FD scenarios (Z1 x_mat = −0.30%), despite modest emission savings (−0.42%), because the food sector is more material-intensive than emissions-intensive.
- **Household energy (Sc2)** produces strong emission and primary material savings (Z1 emis = −0.78%, Z1 x_mat = −0.64%) disproportionate to its channel size (6.9 B EUR) because the 3.3:1 fossil-to-renewable intensity differential amplifies the scale effect on both emission and primary energy extraction.
- **Government plastics (Sc4)** has disproportionate leverage among FD scenarios because the government procurement channel (11.9 B EUR) rivals the intermediate channel in absolute scale — unique among all materials. However, the material effect is ambiguous: EU total material use barely changes (+0.01%) while primary extraction falls only −0.04%, reflecting small primary–secondary intensity gaps for plastics.
- **Energy production (Sc13)** dominates all production scenarios: Z1 x_mat = −1.27% LT (the largest primary material extraction reduction in the entire set), driven by the combination of the largest intermediate channel (14.8 B EUR) and the 3.3:1 intensity differential. Total material use (Z1 mat = −0.04%) moves little because energy restructuring primarily displaces fossil *fuel* use (captured in x_mat) rather than bulk material throughput.
- **Metals production (Sc10)** produces the largest total material savings (Z1 mat = −0.76% LT) because intermediate metal demand (20.9 B EUR) is the largest single intermediate material flow in the scenario set. The A-matrix row-sum of secondary metals (0.238) is less than half that of primary metals (0.545), amplifying material throughput reductions across the supply chain.

The simulation ranking therefore validates the demand channel decomposition: before running the model, relative outcome magnitudes can be predicted from channel scale × intensity differential with reasonable accuracy.

## Two Transition Regimes

The 14 scenarios separate into two regimes distinguished by ecological leverage and macroeconomic direction.

**Final-demand reallocation (Scenarios 1–6)** is not a uniformly low-leverage regime. This characterisation — common in single-region IO analysis — understates the scale of several channels once bilateral MRIO demand flows are included. Scenario 1 (Food, HH) produces the largest employment contraction among FD scenarios (Z1 *n* = −0.09% LT) because food demand (23.7 B EUR) is the dominant household material channel. Scenario 2 (Energy, HH) generates ecological gains (emis −0.78%, x_mat −0.64%) disproportionate to its channel size because the 3.3:1 intensity differential amplifies scale into emission and primary material extraction reduction. Scenario 4 (Gov Plastics) is structurally distinctive: government procurement accounts for 48.9% of plastics demand (11.9 B EUR), making public procurement circularity mandates a genuine first-order CE instrument for this material. Scenario 5 (Construction FD) activates a large fixed-investment channel (88.7 B EUR of total construction demand) and is the only FD scenario producing a positive employment effect (+0.03% LT) because renovation is more labour-intensive than new build. Final-demand CE instruments engage channels that are structurally primary for food, construction, and plastics, and cannot be treated as merely symbolic complements to production-side policy.

**Intermediate production restructuring (Scenarios 7–13)** delivers substantially higher ecological gains at the cost of modest macroeconomic and social contractions. Production scenarios consistently produce larger material savings than their FD counterparts for the same material, in proportion to the intermediate channel's dominance. Employment effects are negative (−0.01% to −0.02% LT) across all intermediate production scenarios except Sc10 (Metals), and wage shares are marginally compressed in the most intensive scenarios (Sc13). The production restructuring regime confirms the core IO insight: where intermediate demand dominates sector activation, A-matrix interventions are the primary CE instrument.

## The CE Trilemma

The cross-cutting finding is a structural trilemma: ecological improvement trades off against macroeconomic stability and distributional equity in the intermediate production scenarios, while the construction scenarios produce a second, ecological-internal tension between material savings and emission rebound.

**Scenario 4 vs Scenario 9 (Plastics)** demonstrates that the trilemma can be partially bypassed through public procurement. For plastics — uniquely among all materials — government consumption (48.9%, 11.9 B EUR) rivals intermediate demand (44.0%, 10.7 B EUR) in absolute scale, making the two channels complementary rather than hierarchical. Neither scenario alone captures a majority of total plastics demand; effective plastics CE policy requires both instruments simultaneously.

**Scenario 10 (Metals)** is the most analytically anomalous outcome. It is the only production scenario showing expansionary macroeconomic effects (Z1 go = +0.04%, *n* = +0.03% LT), driven by the lower A-matrix row-sum of secondary metals (0.238 vs 0.545): substituting secondary for primary frees EU intermediate resources economy-wide and generates a competitive dividend. Simultaneously, emissions increase (+0.27% LT) — the clearest structural rebound in the set. The mechanism is precisely identified: secondary metal production (EAF scrap-based steel) is more electricity-intensive upstream than primary production (blast furnace), and the high fossil emission intensity amplifies this into an EU emission increase even as material throughput falls strongly (Z1 mat = −0.76%). Metals CE satisfies the economic arm of the trilemma while failing the ecological arm and requires complementary electricity decarbonisation.

**Scenario 13 (Energy)** represents maximum leverage and maximum asymmetry simultaneously. Ecological gains are the largest in the set (emis = −1.56%, x_mat = −1.27% LT), and the current account improvement (+1.81%) reflects reduced EU fossil import dependency. But the scenario is distributionally regressive (wage share −0.08%, profit share +0.14%) and produces the largest cross-border burdens on RoW energy exporters (Z2 b_s rises, Z2 *n* = −0.01%). The land use implication is notable: despite saving the most emissions, EU land use *increases* slightly (+0.16% LT) because renewable electricity infrastructure (wind and solar) requires substantially more land per unit of output than centralised fossil power plants. EU energy CE is the highest-leverage domestic CE intervention while simultaneously the most asymmetric in its international consequences.

**Construction scenarios (Sc5, Sc14)** define a third internal tension: material use falls (−0.84% and −0.35% LT respectively) while emissions, land use, and water use all rebound sharply (+5.19% and +2.15%). Renovation/refurbishment is more energy-intensive per unit of output than new construction in the current A-matrix, meaning the circular reallocation displaces primary material throughput but increases upstream fossil energy demand. The ecological trilemma here is not social vs. ecological but rather material vs. atmospheric: circular construction saves physical material but worsens the emission, land, and water footprint unless accompanied by concurrent renovation energy standards and electricity decarbonisation.

## Cross-Border Transmission: Material Flows and RoW Consequences

A key contribution of the MRIO-SFC framework is tracing bilateral demand flows between EU (Z1) and RoW (Z2), making the cross-border consequences of CE policy directly visible. The table below reports long-term RoW material use deviations:

```{r row-mat-table}
mat_vars <- c("Z2_mat", "Z2_x_mat")
mat_lt <- sapply(loads, function(run) pct_vec(run, mat_vars, t_lt))
rownames(mat_lt) <- c("RoW total material use", "RoW primary material extraction")
colnames(mat_lt) <- paste0("Sc", seq_along(loads))
knitr::kable(round(mat_lt, 2), booktabs = TRUE,
             caption = "RoW long-term % deviation from baseline — material use")
```

Four distinct cross-border transmission patterns emerge across the 14 scenarios.

**Symmetric material contraction (most intermediate production scenarios).** Scenarios 7–9, 11–12 produce broadly parallel reductions in material use in both Z1 and Z2, reflecting the global IO linkages through which EU production restructuring reduces demand for RoW intermediate inputs. The effects are small in absolute terms (Z2 mat near zero in most cases) but consistently negative: EU circular restructuring does not simply displace material throughput to the RoW — it reduces it globally. These scenarios are not unilaterally regressive on material grounds.

**Metals (Sc10): the largest RoW material contraction.** The most substantial cross-border material impact occurs in Scenario 10: Z2 total material use falls −0.15% LT and Z2 primary material extraction falls −0.10% LT. This is the largest RoW material reduction in the scenario set — larger, in absolute RoW terms, than any other production scenario. The mechanism is the competitive displacement of RoW metal production: as EU industry shifts to lower-input secondary metals, demand for RoW primary metal exports contracts, pulling down RoW extraction alongside EU extraction. The asymmetry is that RoW loses production and employment while EU macro-expands — a genuine welfare divergence mediated through the material trade channel.

**Plastics and food scenarios: material leakage to RoW.** Scenarios 4 (Gov Plastics) and 9 (Plastics Int) produce small positive RoW material effects (+0.01% mat), indicating partial production leakage: EU secondary plastics production has high import content, and the substitution generates demand for RoW inputs that partially offsets domestic primary material savings. Similarly, Scenario 1 (Food/Meat) produces a marginal RoW material increase (+0.01%), reflecting the displacement of some meat processing toward imported food products. These are the only scenarios in the set where RoW total material use *increases* — a leakage pattern that limits the global material savings of these CE interventions.

**Energy scenarios: RoW primary extraction contraction.** Scenarios 2 and 13 produce RoW primary material extraction contractions (Z2 x_mat = −0.03% and −0.05% LT respectively), driven by the collapse in EU fossil fuel imports. This is the only transmission channel where a large RoW material reduction accompanies a large EU ecological gain. However, this contraction falls asymmetrically on fossil-fuel-exporting economies: the RoW aggregate obscures that specific exporting countries face much larger contractions than the aggregate suggests, with associated employment and fiscal consequences.

**Construction (Sc5, Sc14): RoW material savings despite EU emission rebound.** Despite the emission and land rebound in the EU, both construction scenarios produce small RoW material use *reductions* (Z2 mat = −0.05% in Sc5, −0.02% in Sc14) because renovation uses less imported primary construction material than new build. The cross-border material effect is therefore mildly positive globally even when the EU domestic ecological balance is mixed.

### Ecological Pressure Divergence: Land and Water

A secondary but analytically important finding is that emissions, land use, and water use diverge across scenarios. The three pressure dimensions are driven by different sector-specific intensity coefficients:

- **Scenario 1 (Food/Meat):** Emissions fall −0.42% but water use *increases* +0.68% EU — other food products (the secondary sector) are more water-intensive per unit of output than the primary meat sector. Land use falls only −0.12%, much less than emissions.
- **Scenario 3 (Wood HH):** Emissions are near zero but land use falls −0.36% — primary wood production is highly land-intensive; the circular shift releases land without changing fossil energy consumption.
- **Scenario 7 (Wood Int):** The **largest land saving in the entire scenario set** (EU land = −2.30% LT) from a scenario with near-zero emission effect (−0.01%). The land–emission decoupling is strongest here: forest-sector restructuring delivers a land dividend that the emission indicator entirely misses.
- **Scenario 13 (Energy Int):** Emissions fall −1.56% but land use *increases* +0.16% — renewable energy infrastructure requires more land per MWh than fossil plants. The energy scenario is simultaneously the best CE intervention for climate and the worst for land use.
- **Construction (Sc5, Sc14):** Emissions rebound (+5.19%, +2.15%) but land and water *improve* (−0.33%, −0.05% in Sc5). The circular reallocation reduces primary material and land extraction while worsening the atmospheric footprint.

These divergences reinforce that no single ecological indicator is sufficient for CE policy evaluation. Wood-sector restructuring is primarily a land-use intervention; energy-sector restructuring is primarily a climate intervention; food-sector restructuring is primarily an atmospheric and land intervention that worsens freshwater pressure.

## Policy Implications

Four policy implications follow directly from the simulation evidence.

First, **production restructuring requires macroeconomic stabilisation.** Production scenarios consistently contract aggregate demand modestly, reduce wage shares, and expand profit shares. CE transitions in the production regime cannot be treated as costless structural adjustments — they require active macroeconomic management and labour market support.

Second, **final-demand reallocation is primary for specific channels.** For most material sectors where intermediate demand dominates (metals, wood, glass, cement), FD instruments engage structurally secondary channels and cannot substitute for A-matrix restructuring. However, for construction (investment-dominated), food (household-dominated), and plastics (government-procurement-dominated), FD instruments engage the primary demand channel directly and are first-order CE instruments in their own right.

Third, **energy is the highest-leverage CE instrument, but land use is its shadow cost.** Scenario 13 dominates all other production scenarios in both emission reduction (−1.56%) and primary material extraction reduction (−1.27%). However, renewable energy infrastructure requires substantially more land per unit of output than fossil plants, meaning that scaling this transition without land-use planning creates a new ecological trade-off that the emissions indicator does not capture.

Fourth, **metals CE requires electricity decarbonisation as a complement, and its cross-border material dividend is real but regressive.** Without simultaneous electricity decarbonisation, secondary metal production amplifies upstream emissions through the fossil energy link. With decarbonisation, Metals CE could simultaneously expand EU output, reduce global material throughput (including −0.15% in RoW), and reduce emissions — making it uniquely triple-positive. The current calibration shows only the first and second arm; the third requires a policy complement.

## Validation of the Descriptive Analytics

The simulation ranking matches the demand channel ranking predicted by the portfolio decomposition. Among FD scenarios, the ordering Food (mat −0.51%) > Energy (x_mat −0.64%) > Plastics (near-neutral material) > Wood (negligible) is consistent with the household and government channel scales weighted by intensity differentials. Among production scenarios, Energy > Metals (material) > Wood > Glass > Cement > Pulp is consistent with the intermediate channel scales and intensity differentials. The descriptive analytics' prediction that portfolio scale — not material-specific multiplier variation — determines the simulation ordering is confirmed across all 14 scenarios.
