---
title: "Effective climate action can only be global in scope"
subtitle: "Insights from a Core-Periphery SOCIO-ECO-MRIO-SFC Model"
author:
  - name: "Oriol Vallès Codina"
    affiliation: "Johns Hopkins University, Net Zero Industrial Policy Lab, Department of Political Science"
    email: 
      - "oriolvallescodina@gmail.com"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc_float: true
    self-contained: true
    page-layout: full
    css: fullwidth.css
execute:
  echo: false
---

# Initialization

The initial values and variables are loaded from the file stored in variable `identif` in a different structure as before.

```{r packages, warning = F, message = F}
library(tidyverse)
library(igraph)
library(qgraph)
library(visNetwork)
library(scales)
```

```{r init, warning = FALSE, message = FALSE}
project_root <- rprojroot::find_root(
  rprojroot::is_rstudio_project | rprojroot::is_git_root
)

variable.table <<- read.csv(
  file.path(project_root, "data", "Variable_Definitions.csv")
)

source(file.path(project_root, "functions", "auxiliary_flexible_nov25.R"))

identif <- file.path(project_root, "data", "initial_state_2026.xlsx")
initial <- load.new.init(identif)

source(file.path(project_root, "functions", "run_model_2026.R"))
source(file.path(project_root, "functions", "MVP_model_2026.R"))
```

# Baseline 

## Simulation

```{r simulation, warning = F, message = F}
baseline <- run.model(
  initial,
  mvp.model,
  log_file = file.path(project_root, "logs", "baseline_run.log"),
  log_append = FALSE,
  print_final_state = FALSE
)

saveRDS(baseline, file = file.path(project_root, 'data', 'baseline_feb10.RDS'))
```

## Calibration

Target values are loaded in format long (`target.set`) and wide (`target.table.0`). Units are divided by 10,000 for convenience.

```{r calibration}
#### LOAD TARGET VALUES ####

target.list <- load.target(file.path(
  project_root,
  'data',
  'target_values.xlsx'
))
target.table <- target.list$wide
target.set <- target.list$long
target.vars <- target.set$var.label
target.set$target <- baseline$simulation[
  target.set$var,
  baseline$initial$pars['nPeriods', 'value']
]
target.set$ratio <- round(target.set$target / target.set$value, 2)

compute.target(baseline)
```

## Visualization

```{r variable_selection} 
# Variable Selection
selected.list <- list(
  Macroeconomic = c('c', 'rva', 'va', 'go', 'id', 'gdef'),
  External = c('cab', 'tb'),
  Social = c('n', 'nf', 'shp', 'shw'),
  # Prices = c('pa', 'pid', 'pg', 'pim'),
  Employment = c('n', 'nf'),
  Inequality = c('ydw', 'ydc', 'shp', 'shw'),
  Debt.and.Wealth = c('lh', 'lf', 'v', 'k', 'b_s'),
  Ecological = c('x_mat', 'land', 'mat', 'rec', 'emis', 'wa', 'water')
)
```

```{r plot_by_category}
plot.selected.by_group(baseline, selected.list)
```

# Scenario Analysis

| Domain      | Shift              | From | To | Mechanism                  | Conserved Quantity            |
|-------------|--------------------|------|----|----------------------------|-------------------------------|
| Household   | Diet shift         | 7    | 8  | Final demand reallocation  | Total food expenditure        |
| Household   | Energy transition  | 31   | 32 | Final demand reallocation  | Energy service demand         |
| Production  | Wood circularity   | 11   | 12 | Technical coefficient change | Total wood material input   |
| Production  | Pulp circularity   | 13   | 14 | Technical coefficient change | Paper output                |
| Production  | Plastics circularity | 17 | 18 | Technical coefficient change | Plastic product output      |
| Production  | Metal circularity  | 26   | 27 | Technical coefficient change | Metal product output        |
| Production  | Glass circularity  | 21   | 22 | Technical coefficient change | Glass product output        |
| Production  | Cement circularity | 24   | 25 | Technical coefficient change | Cementitious output         |
| Production  | Energy transition  | 31   | 32 | Technical coefficient change | Energy service to production|
| Production  | Construction reuse | 36   | 27 | Technical coefficient change | Construction output         |

## Diet Shift

```{r shock.run}
# Time Period Selection
t.shock <- initial$pars['t.shock', 'value'] # shock time period
t0 <- 55 # initial time period for visualization
tf <- initial$pars['nPeriods', 'value'] # final time period for visualization

# Shock Selection and Run
initial_shock <- initial
initial_shock$pars['shock', 'value'] <- 1

diet.shock <- run.model(
  initial_shock,
  mvp.model,
  log_file = file.path(project_root, "logs", "shock_1_run.log"),
  log_append = FALSE,
  print_final_state = FALSE
)
saveRDS(diet.shock, file = 'data/shock_1_diet_shock.RDS')
# ple.shock <- readRDS(file = 'data/pleshock_aug14.RDS')

shock.title <- 'Scenario: Diet Shift'
```

## Visualize Shocked Variables

```{r shock.view}
view.shock(
  diet.shock,
  shock.vars = c(
    'delta',
    'id0',
    'iota_g-1',
    'iota_g-46',
    'zeta_dc-28',
    'beta-28',
    'beta-29',
    'beta-41'
  ),
  shock.sectors = c(NA, NA, NA, 46, 28, 28, 29, 41)
)
```

## Visualize Impacts

```{r shock.visualize.scaled}

# Load Baseline - Shock Data
df <- shock.long.new(baseline, diet.shock)

# Plot Selected List Variables (scaled only if gdef is not there)
for (i in 1:length(selected.list)) {
  if ('gdef' %in% selected.list[[i]]) {
    print(view.vars(
      data = df,
      viz.vars = selected.list[[i]],
      var.label = names(selected.list)[i]
    ))
  } else {
    print(view.scaled.vars(
      data = df,
      viz.vars = selected.list[[i]],
      var.label = names(selected.list)[i]
    ))
  }
}

# Print Some Scaled Variables Example
# view.scaled.vars(data = shock.long.new(baseline, ple.shock),
#                  viz.vars = c('cab', 'ydw', 'n', 'rva'),
#                  var.label = 'Scaled')

# Print Some Variables Example (not scaled)
# view.vars(data = shock.long.new(baseline, ple.shock),
#           viz.vars = c('cab', 'ydw', 'n', 'rva'),
#           var.label = 'Not Scaled')
```

## Visualize Selected Variables Without Function

```{r shock.visualize.selected.2}
# Selected Vars to plot
selected <- c('n', 'c', 'ineq', 'va', 'cab', 'nf', 'gdef', 'tb', 'go')

p.selected.vars <- filter(df, variable %in% selected) %>%
  ggplot(aes(x = time, y = value, color = area, linetype = scenario)) +
  facet_wrap(~name, scales = 'free') +
  geom_line() +
  geom_vline(
    xintercept = initial$pars['t.shock', 'value'],
    linetype = 'dashed',
    linewidth = .4
  ) +
  labs(
    title = shock.title,
    subtitle = 'Selected Indicators. Vertical dashed line indicates shock time'
  )

p.selected.vars
```


