---
title: "Scenario Analysis"
subtitle: "Insights from a Core-Periphery SOCIO-ECO-MRIO-SFC Model"
author:
  - name: "Oriol Vallès Codina"
    affiliation: "Johns Hopkins University, Net Zero Industrial Policy Lab, Department of Political Science"
    email:
      - "oriolvallescodina@gmail.com"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc_float: true
    self-contained: true
    page-layout: full
    css: fullwidth.css
execute:
  echo: false
  working-dir: ..
---

# Initialization

The initial values and variables are loaded from the file stored in `initial_state_2026.xlsx`. This file is the **analysis layer**: it reads cached runs from `output/scenarios/` and produces plots and interpretation.

```{r packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(igraph)
library(qgraph)
library(visNetwork)
library(scales)
library(knitr)
library(kableExtra)
```

```{r init, warning=FALSE, message=FALSE}
# -----------------------------
# Robust project root detection
# -----------------------------
qmd_in <- knitr::current_input()

if (!is.null(qmd_in) && nzchar(qmd_in)) {
  qmd_path <- normalizePath(qmd_in, winslash = "/")
  root <- normalizePath(file.path(dirname(qmd_path), ".."), winslash = "/")
  root_method <- "knitr::current_input()"
} else {
  wd0 <- normalizePath(getwd(), winslash = "/")
  root <- if (basename(wd0) == "qmd") normalizePath("..", winslash = "/") else wd0
  root_method <- "getwd() fallback"
}

dir_data <- file.path(root, "data")

cat("root_method:", root_method, "\n")
cat("getwd():", normalizePath(getwd(), winslash = "/"), "\n")
cat("root  :", root, "\n")
cat("dir_data:", dir_data, "\n")

# Canonical directories
model_dir <- file.path(root, "model")

# Core inputs
variable.table <- read.csv(
  file.path(dir_data, "Variable_Definitions.csv"),
  stringsAsFactors = FALSE
)

initial_filename <- file.path(dir_data, "initial_state_2026.xlsx")
stopifnot(file.exists(initial_filename))

sector_list <- read.csv(
  file.path(dir_data, "sector_list.csv"),
  stringsAsFactors = FALSE
) %>%
  mutate(label = gsub("<br>", " ", label))

# Workspace dirs
workspace_dir <- file.path(root, "output", "scenarios")
dir_runs  <- file.path(workspace_dir, "shock_runs")
dir_tables <- file.path(workspace_dir, "shock_tables")
dir_logs  <- file.path(workspace_dir, "logs")
dir_long  <- file.path(workspace_dir, "shock_long")
dir_cache <- file.path(workspace_dir, "scenario_cache")
dir_rds   <- file.path(workspace_dir, "rds_files")

# Workspace filenames
baseline_filename <- file.path(workspace_dir, "baseline_2026.RDS")
target_filename <- file.path(workspace_dir, "target_values.xlsx")
scenario_filename <- file.path(workspace_dir, "scenarios.csv")

stopifnot(file.exists(baseline_filename))
stopifnot(file.exists(target_filename))
stopifnot(file.exists(scenario_filename))

sc <- read.csv(file = scenario_filename, stringsAsFactors = FALSE)

get_shock_filename <- function(n) file.path(dir_runs, paste0("shock_", n, "_run.RDS"))
get_table_filename <- function(n) file.path(dir_tables, paste0("shock_", n, "_table.csv"))
get_log_filename   <- function(n) file.path(dir_logs, paste0("shock_", n, "_run.log"))

# Source model + helpers
source(file.path(model_dir, "bootstrap_2026.R"))

# Load initial state (for parameters like t.shock and nPeriods)
initial <- load.init(initial_filename)
```

```{r theme, warning=FALSE, message=FALSE}
theme_leeds <- function() {
  theme_grey() +
    theme(
      plot.title = element_text(face = "bold", size = 14),
      legend.title = element_text(face = "bold"),
      strip.text.y = element_text(angle = 0),
      plot.subtitle = element_text(face = "italic", size = 10),
      strip.text.y.right = element_text(angle = 0)
    )
}
theme_set(theme_leeds())
```

# Baseline

```{r baseline, warning=FALSE, message=FALSE}
baseline <- readRDS(file = baseline_filename)
```

## Calibration

```{r calibration, warning=FALSE, message=FALSE}
target.list <- load.target(target_filename, variable.table)
target.table <- target.list$wide
target.set <- target.list$long
target.vars <- target.set$var.label

target.set$target <- baseline$simulation[
  target.set$var,
  baseline$initial$pars["nPeriods", "value"]
]
target.set$ratio <- round(target.set$target / target.set$value, 2)

compute.target(baseline)
```

## Visualization

```{r variable_selection}
#| fig-width: 7
#| fig-height: 4

selected.list <- list(
  Macroeconomic = c("c", "va", "go", "id", "gdef", "cab", "tb"),
  Social = c("n", "nf", "shp", "shw"),
  Debt.and.Wealth = c("lh", "lf", "v", "k", "b_s"),
  Ecological = c("x_mat", "land", "mat", "rec", "emis", "wa", "water")
)

plot.selected.by_group(baseline, selected.list)
```

---

# Scenarios to Analyze

```{r scenarios_table, warning=FALSE, message=FALSE}
sc %>%
  select(-any_of("display_name")) %>%
  kable(
    caption = "Scenario Registry",
    align = "l",
    escape = TRUE
  ) %>%
  kable_styling(
    full_width = FALSE,
    bootstrap_options = c("striped", "hover", "condensed"),
    font_size = 12
  ) %>%
  scroll_box(width = "100%", height = "600px")

# Time Period Selection
t.shock <- initial$pars["t.shock", "value"]
t0 <- 55
tf <- initial$pars["nPeriods", "value"]
```

# Scenario results

This section assumes scenario outputs have been cached as `output/scenarios/shock_runs/shock_<id>_run.RDS`.

```{r load_shocks, warning=FALSE, message=FALSE}
shock_ids <- sort(unique(as.integer(sc$shock)))
available <- shock_ids[file.exists(vapply(shock_ids, get_shock_filename, character(1)))]

cat("Available cached shocks:", paste(available, collapse = ", "), "\n")

shock_list <- setNames(
  lapply(available, function(s) readRDS(get_shock_filename(s))),
  paste0("shock_", available)
)
```

## View shocked variables

If your utilities define `view.shock()`, this reproduces the *shocked-variable* diagnostics without re-running the model.

```{r view_shock, warning=FALSE, message=FALSE}
if (exists("view.shock", mode = "function")) {
  # By default show the first two cached shocks (edit as needed)
  show_ids <- head(available, 2)

  for (s in show_ids) {
    cat("\n\n## Shock ", s, "\n\n", sep = "")
    view.shock(
      shock = shock_list[[paste0("shock_", s)]],
      baseline = baseline,
      sc = sc,
      shock_id = s,
      t0 = t0,
      tf = tf
    )
  }
} else {
  cat("view.shock() not found. Define it (or source its file) in model/bootstrap_2026.R.\n")
}
```
