---
title: "Scenario Design"
subtitle: "Insights from a Core-Periphery SOCIO-ECO-MRIO-SFC Model"
author:
  - name: "Oriol Vallès Codina"
    affiliation: "Johns Hopkins University, Net Zero Industrial Policy Lab, Department of Political Science"
    email:
      - "oriolvallescodina@gmail.com"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc_float: true
    self-contained: true
    number-sections: true
    page-layout: full
    css: fullwidth.css
execute:
  echo: false
  working-dir: ..
---

# Initialization

This document defines the scenario *registry* and the *formal* intervention operators. It is intentionally non-executing: it loads the baseline and metadata, but does **not** run shocks.

```{r packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(igraph)
library(qgraph)
library(visNetwork)
library(scales)
library(knitr)
library(kableExtra)
```

```{r init, warning=FALSE, message=FALSE}
# -----------------------------
# Robust project root detection
# -----------------------------
qmd_in <- knitr::current_input()

if (!is.null(qmd_in) && nzchar(qmd_in)) {
  qmd_path <- normalizePath(qmd_in, winslash = "/")
  root <- normalizePath(file.path(dirname(qmd_path), ".."), winslash = "/")
  root_method <- "knitr::current_input()"
} else {
  # Fallback for interactive execution / sourcing
  wd0 <- normalizePath(getwd(), winslash = "/")
  root <- if (basename(wd0) == "qmd") {
    normalizePath("..", winslash = "/")
  } else {
    wd0
  }
  root_method <- "getwd() fallback"
}

dir_data <- file.path(root, "data")

cat("root_method:", root_method, "\n")
cat("getwd():", normalizePath(getwd(), winslash = "/"), "\n")
cat("root  :", root, "\n")
cat("dir_data:", dir_data, "\n")

# -----------------------------
# Canonical directories
# -----------------------------
model_dir <- file.path(root, "model")
functions_dir <- file.path(root, "functions") # keep only if you still use it
utils_dir <- file.path(root, "utils")

# -----------------------------
# Core inputs
# -----------------------------
variable.table <- read.csv(
  file.path(dir_data, "Variable_Definitions.csv"),
  stringsAsFactors = FALSE
)

initial_filename <- file.path(dir_data, "initial_state_2026.xlsx")
stopifnot(file.exists(initial_filename))

sector_list <- read.csv(
  file.path(dir_data, "sector_list.csv"),
  stringsAsFactors = FALSE
) %>%
  mutate(label = gsub("<br>", " ", label))

# -----------------------------
# Workspace dirs
# -----------------------------
workspace_dir <- file.path(root, "output", "scenarios")
dir.create(workspace_dir, showWarnings = FALSE, recursive = TRUE)

dir_runs <- file.path(workspace_dir, "shock_runs")
dir_tables <- file.path(workspace_dir, "shock_tables")
dir_logs <- file.path(workspace_dir, "logs")
dir_long <- file.path(workspace_dir, "shock_long")
dir_cache <- file.path(workspace_dir, "scenario_cache")
dir_rds <- file.path(workspace_dir, "rds_files")

invisible(lapply(
  list(dir_runs, dir_tables, dir_logs, dir_long, dir_cache, dir_rds),
  dir.create,
  showWarnings = FALSE,
  recursive = TRUE
))

# -----------------------------
# Source utilities + model code
# -----------------------------
# Assumed to define:
# - load.init()
# - run.model()
# - mvp.model (or equivalent)
# - load.target(), compute.target(), plotting helpers
# - (optionally) compute_scenario(), view.shock(), etc.
source(file.path(model_dir, "bootstrap_2026.R"))

# --- Load initial state ---
initial <- load.init(initial_filename)
```

```{r filenames, warning=FALSE, message=FALSE}
# ---- Workspace filenames (generated / cached) ----
baseline_filename <- file.path(workspace_dir, "baseline_2026.RDS")
target_filename <- file.path(workspace_dir, "target_values.xlsx")
scenario_filename <- file.path(workspace_dir, "scenarios.csv")

stopifnot(file.exists(baseline_filename))
stopifnot(file.exists(target_filename))
stopifnot(file.exists(scenario_filename))

sc <- read.csv(file = scenario_filename, stringsAsFactors = FALSE)

get_shock_filename <- function(n) {
  file.path(dir_runs, paste0("shock_", n, "_run.RDS"))
}
get_table_filename <- function(n) {
  file.path(dir_tables, paste0("shock_", n, "_table.csv"))
}
get_log_filename <- function(n) {
  file.path(dir_logs, paste0("shock_", n, "_run.log"))
}
```

# Baseline

```{r baseline, warning=FALSE, message=FALSE}
baseline <- readRDS(file = baseline_filename)
```

# Calibration targets

Target values are loaded in format long (`target.set`) and wide (`target.table`). The baseline-implied values are copied into `target.set$target` to enable quick ratio checks.

```{r calibration, warning=FALSE, message=FALSE}
target.list <- load.target(target_filename, variable.table)
target.table <- target.list$wide
target.set <- target.list$long

target.set$target <- baseline$simulation[
  target.set$var,
  baseline$initial$pars["nPeriods", "value"]
]
target.set$ratio <- round(target.set$target / target.set$value, 2)

compute.target(baseline)
```

# Scenario registry

This is the *single* authoritative list of scenarios. Every analysis or run script should treat `scenarios.csv` as the source of truth.

```{r scenarios_table, warning=FALSE, message=FALSE}
sc %>%
  kable(
    caption = "Scenario Registry",
    align = "l",
    escape = TRUE
  ) %>%
  kable_styling(
    full_width = FALSE,
    bootstrap_options = c("striped", "hover", "condensed"),
    font_size = 12
  ) %>%
  scroll_box(width = "100%", height = "600px")
```

# Formal intervention operators

## Objects and accounting identities

Let $A \in \mathbb{R}^{N\times N}$ be the technical coefficients matrix and $\mathbf{d}\in\mathbb{R}^N$ the real final demand vector. Output is:

$$
\mathbf{x} = (I-A)^{-1}\mathbf{d} \equiv L\mathbf{d}.
$$

Let material sectors of interest be indexed by $i\in\mathcal{M}$ (in the simplest case $\mathcal{M}=\{1,2\}$). Sectoral material use is:

$$
M_i = x_i, \qquad M = \sum_{i\in\mathcal{M}} M_i.
$$

We consider two **distinct** scenario classes:

1. **Final-demand reallocation**: change the *composition* of a demand component while holding its total level fixed.
2. **Intermediate-input substitution**: change *columns* of $A$ while conserving an input bundle within each using sector.

Mixing these two without saying so is a conceptual error: they operate through different accounting channels and generate different multiplier logic.

## Final-demand reallocation

Let $D \in \{c,g,id,id^g\}$ denote the level of a demand component and let $\delta\in\{\beta,\sigma,\iota,\iota^g\}$ denote its sector-level shares so that:

$$
\sum_{j=1}^N \delta_j = 1, \qquad
\mathbf{d}^{(D)} = D\,\delta.
$$

A directed reallocation from sector $k$ to sector $l$ with intensity $\rho\in[0,1]$ is:

$$
\delta'_k = (1-\rho)\,\delta_k,\qquad
\delta'_l = \delta_l + \rho\,\delta_k,\qquad
\delta'_j = \delta_j \;\; (j\notin\{k,l\}).
$$

This preserves the simplex:

$$
\sum_j \delta'_j = \sum_j \delta_j = 1,
\qquad
\sum_j (\delta'_j-\delta_j) = 0.
$$

The corresponding change in demand is:

$$
\Delta \mathbf{d}^{(D)} = D\,(\delta' - \delta).
$$

With fixed $A$, output responds as:

$$
\Delta \mathbf{x} = L\,\Delta\mathbf{d}^{(D)}.
$$

### What $\rho$ is (and what it is not)

$\rho$ is *not* a productivity parameter, and it does not alter the level $D$. It is purely a **composition** parameter: the fraction of sector-$k$'s share that is reassigned to sector $l$ within the relevant demand component.

If you want a scenario where total demand changes (e.g., absolute investment increases), that is a different operator and should be represented explicitly as $D'\neq D$.

## Intermediate-input substitution

Let sectors $1$ (primary) and $2$ (secondary) supply the same functional material input into every using sector $j$. For each $j$, define the input bundle:

$$
s_j \equiv a_{1j}+a_{2j}.
$$

A substitution from primary to secondary at rate $\rho\in[0,1]$ is defined column-wise:

$$
a'_{1j} = (1-\rho)\,a_{1j},\qquad
a'_{2j} = a_{2j} + \rho\,a_{1j},\qquad
a'_{ij} = a_{ij} \;\; (i\neq 1,2).
$$

Bundle conservation is explicit:

$$
a'_{1j}+a'_{2j} = a_{1j}+a_{2j} = s_j.
$$

The exact output under $A'$ is:

$$
\mathbf{x}' = (I-A')^{-1}\mathbf{d}.
$$

For small changes (useful for intuition, not as a substitute for simulation), the first-order approximation is:

$$
\Delta\mathbf{x} \approx L\,(\Delta A)\,L\,\mathbf{d},
\quad \text{with } \Delta A = A'-A.
$$

## Scenario table requirements

To make the pipeline reproducible, every row of `scenarios.csv` must minimally encode:

- `shock` (integer id)
- `kind` in `{final_demand, intermediate}`
- `component` in `{c,g,id,id_g,A}` (or equivalent naming)
- `from` (sector id $k$)
- `to` (sector id $l$) — only relevant for final-demand reallocations
- `rho` in $[0,1]$
- optional: `t_shock`, `notes`, `prefix` (if your variable naming scheme needs it)

If the table does not contain these fields (or their clear equivalents), your run scripts will necessarily become ad hoc and error-prone.
