---
title: "Effective climate action can only be global in scope"
subtitle: "Insights from a Core-Periphery SOCIO-ECO-MRIO-SFC Model"
author:
  - name: "Oriol Vallès Codina"
    affiliation: "Johns Hopkins University, Net Zero Industrial Policy Lab, Department of Political Science"
    email: 
      - "oriolvallescodina@gmail.com"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc_float: true
    self-contained: true
    page-layout: full
    css: fullwidth.css
execute:
  echo: false
---

# Initialization

The initial values and variables are loaded from the file stored in variable `identif` in a different structure as before.

```{r init, warning = FALSE, message = FALSE}
# Clear Environment
rm(list = ls())

# Load Auxiliary Functions
# directory <- '' # '' for the markdown file, 'flexible_/' for R
variable.table <<- read.csv('data/Variable_Definitions.csv')
# source('functions/auxiliary_flexible_aug12.R') # load.new.init, auxiliary labeling functions, plotting
source('functions/auxiliary_flexible_feb7.R')


# Load Initial Values
identif <- 'data/perfect_init_aug14_withBmatrix_PLE_CE.xlsx'
initial <- load.new.init(identif)

# Load Models
source("functions/run_new_model_feb7.R") # function run.new.model(initial.values, model.equations)
source("functions/MVP_model_aug14.R") # model equations

# baseline <- run.new.model(initial, mvp.model)
# saveRDS(baseline, file = 'data/baseline_aug14.RDS')
baseline <- readRDS(file = 'data/baseline_aug14.RDS')
```

# Table of Industries

```{r industry.table}
kable(
  data.frame(number = 1:length(initial$sectors), industry = initial$sectors),
  booktabs = T
)
```


# MRIO $A$ Matrix Visualization

Importantly, the column names of the $A$ matrix contained in `identif` determine the country and industry labels in the whole model, following the grammar `country_label.industry_label`:

```{r network}
library(igraph)
library(qgraph)
go <- rowSums(initial$A.matrix)
g <- graph_from_adjacency_matrix(
  as.matrix(initial$A.matrix),
  mode = 'directed',
  weighted = TRUE,
  diag = F,
  add.colnames = T
)

# Graph Visualization
l <- qgraph.layout.fruchtermanreingold(
  get.edgelist(g, names = FALSE),
  vcount = vcount(g)
)
V(g)$size <- 7.5 * go^0.75
V(g)$frame.color <- NA
# V(g)$name <- apply(expand.grid(1 : unique(initial$K), initial$countries)[c(2,1)], 1, function(x) paste0(x, collapse = ""))
V(g)$name <- rep(1:unique(initial$K), times = 2)
V(g)$color <- rep(scales::hue_pal()(4)[c(1, 3)], each = unique(initial$K))
E(g)$color <- alpha('black', E(g)$weight / max(E(g)$weight))

par(mar = c(0, 0, 0, 0))
plot(g, edge.arrow.size = 0)
# title("MRIO (EU, red; RoW, emerald)")
```

# Calibration

Target values are loaded in format long (`target.set`) and wide (`target.table.0`). Units are divided by 10,000 for convenience.

```{r target}
#### LOAD TARGET VALUES ####

target.list <- load.target('data/target_values.xlsx')
target.table <- target.list$wide
target.set <- target.list$long
target.vars <- target.set$var.label
target.set$target <- baseline$simulation[
  target.set$var,
  baseline$initial$pars['nPeriods', 'value']
]
target.set$ratio <- round(target.set$target / target.set$value, 2)

compute.target(baseline)
```

# Shock Example 7: Product Lifetime Extension

## Run Shock

This is a simple visualization of shock 7. Please note the new grammar to change initial values.

```{r shock.run}
# Time Period Selection
t.shock <- initial$pars['t.shock', 'value'] # shock time period
t0 <- 55 # initial time period for visualization
tf <- initial$pars['nPeriods', 'value'] # final time period for visualization

# Shock Selection and Run
initial$pars['shock', 'value'] <- 7 # shock 7
# ple.shock <- run.new.model(initial, mvp.model)
# saveRDS(ple.shock, file = 'data/pleshock_aug14.RDS')
ple.shock <- readRDS(file = 'data/pleshock_aug14.RDS')

shock.title <- 'Scenario: Product Lifetime Extension'
```

## Visualize Shocked Variables

```{r shock.view}
view.shock(
  ple.shock,
  shock.vars = c(
    'delta',
    'id0',
    'iota_g-1',
    'iota_g-46',
    'zeta_dc-28',
    'beta-28',
    'beta-29',
    'beta-41'
  ),
  shock.sectors = c(NA, NA, NA, 46, 28, 28, 29, 41)
)
```

## Visualize Impacts

```{r shock.visualize.scaled}
# Variable Selection
selected.list <- list(
  Economic = c('n', 'c', 'ineq', 'va', 'cab', 'nf', 'gdef', 'tb', 'go'),
  Macroeconomic = c('c', 'rva', 'go', 'id'),
  External = c('cab', 'gdef', 'tb'),
  Social = c('n', 'nf', 'shp', 'shw'),
  Prices = c('pa', 'pid', 'pg', 'pim'),
  Employment = c('n', 'nf'),
  Inequality = c('ydw', 'ydc', 'ineq', 'shp', 'shw'),
  Debt.and.Wealth = c('lh', 'lf', 'v', 'k', 'b_s'),
  Ecological = c('x_mat', 'mat', 'rec', 'emis', 'wa')
)

# Load Baseline - Shock Data
df <- shock.long.new(baseline, ple.shock)

# Plot Selected List Variables (scaled only if gdef is not there)
for (i in 1:length(selected.list)) {
  if ('gdef' %in% selected.list[[i]]) {
    print(view.vars(
      data = df,
      viz.vars = selected.list[[i]],
      var.label = names(selected.list)[i]
    ))
  } else {
    print(view.scaled.vars(
      data = df,
      viz.vars = selected.list[[i]],
      var.label = names(selected.list)[i]
    ))
  }
}

# Print Some Scaled Variables Example
# view.scaled.vars(data = shock.long.new(baseline, ple.shock),
#                  viz.vars = c('cab', 'ydw', 'n', 'rva'),
#                  var.label = 'Scaled')

# Print Some Variables Example (not scaled)
# view.vars(data = shock.long.new(baseline, ple.shock),
#           viz.vars = c('cab', 'ydw', 'n', 'rva'),
#           var.label = 'Not Scaled')
```

## Visualize Selected Variables Without Function

```{r shock.visualize.selected.2}
# Selected Vars to plot
selected <- c('n', 'c', 'ineq', 'va', 'cab', 'nf', 'gdef', 'tb', 'go')

p.selected.vars <- filter(df, variable %in% selected) %>%
  ggplot(aes(x = time, y = value, color = area, linetype = scenario)) +
  facet_wrap(~name, scales = 'free') +
  geom_line() +
  geom_vline(
    xintercept = initial$pars['t.shock', 'value'],
    linetype = 'dashed',
    linewidth = .4
  ) +
  labs(
    title = shock.title,
    subtitle = 'Selected Indicators. Vertical dashed line indicates shock time'
  )

p.selected.vars
```


## Summary Table

```{r ple.summary}
# Summary Table for PLE Scenario
ple.table <- shock.summary(
  baseline,
  shock.run = ple.shock,
  t_ = c(t.shock + 1, t.shock + 7, t.shock + 20),
  t_names = c('Immediate', 'Short.Term', 'Long.Term')
)

kable(ple.table, booktabs = T)
```

# Shock Example 4: Higher Use of Circular Economy Inputs

## Run Shock

This is a simple visualization of shock 4. Please note the new grammar to change initial values.

```{r shock.run.2}
# Shock Selection and Run
init.ce <- load.new.init(
  'data/perfect_init_aug14_withBmatrix_CE_inputs_higher_costs.xlsx'
)

# Time Period Selection
t.shock <- init.ce$pars['t.shock', 'value'] # shock time period
t0 <- 55 # initial time period for visualization
tf <- init.ce$pars['nPeriods', 'value'] # final time period for visualization

init.ce$pars['shock', 'value'] <- 4

# ce.shock <- run.new.model(init.ce, mvp.model)
# saveRDS(ce.shock, file = 'data/CE_shock_aug14.RDS')
ce.shock <- readRDS(file = 'data/CE_shock_aug14.RDS')

shock.title <- 'Scenario: Higher Use of Circular Economy Inputs'
```

## Visualize Shocked Variables

```{r shock.view.2}
view.shock(
  ce.shock,
  shock.vars = c('delta', 'vat-12'),
  shock.sectors = c(NA, 12)
)
```

## Visualize Impacts

```{r shock.visualize.scaled.2}
# Variable Selection
selected.list <- list(
  Economic = c('n', 'c', 'ineq', 'va', 'cab', 'nf', 'gdef', 'tb', 'go'),
  Macroeconomic = c('c', 'rva', 'go', 'id'),
  External = c('cab', 'gdef', 'tb'),
  Social = c('n', 'nf', 'shp', 'shw'),
  Prices = c('pa', 'pid', 'pg', 'pim'),
  Employment = c('n', 'nf'),
  Inequality = c('ydw', 'ydc', 'ineq', 'shp', 'shw'),
  Debt.and.Wealth = c('lh', 'lf', 'v', 'k', 'b_s'),
  Ecological = c('x_mat', 'mat', 'rec', 'emis', 'wa')
)

# Load Baseline - Shock Data
df <- shock.long.new(baseline, ce.shock)

# Plot Selected List Variables (scaled only if gdef is not there)
for (i in 1:length(selected.list)) {
  if ('gdef' %in% selected.list[[i]]) {
    print(view.vars(
      data = df,
      viz.vars = selected.list[[i]],
      var.label = names(selected.list)[i]
    ))
  } else {
    print(view.scaled.vars(
      data = df,
      viz.vars = selected.list[[i]],
      var.label = names(selected.list)[i]
    ))
  }
}

# Print Some Scaled Variables Example
# view.scaled.vars(data = shock.long.new(baseline, ple.shock),
#                  viz.vars = c('cab', 'ydw', 'n', 'rva'),
#                  var.label = 'Scaled')

# Print Some Variables Example (not scaled)
# view.vars(data = shock.long.new(baseline, ple.shock),
#           viz.vars = c('cab', 'ydw', 'n', 'rva'),
#           var.label = 'Not Scaled')
```

## Summary Table

```{r summary.table.2}
ce.table <- shock.summary(
  baseline,
  shock.run = ce.shock,
  t_ = c(t.shock + 1, t.shock + 7, t.shock + 20),
  t_names = c('Immediate', 'Short.Term', 'Long.Term')
)

kable(ce.table, booktabs = T)
```

