---
title: "Scenario Analysis"
subtitle: "Insights from a Core-Periphery SOCIO-ECO-MRIO-SFC Model"
author:
  - name: "Oriol Vallès Codina"
    affiliation: "Johns Hopkins University, Net Zero Industrial Policy Lab, Department of Political Science"
    email: 
      - "oriolvallescodina@gmail.com"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc_float: true
    self-contained: true
    page-layout: full
    css: fullwidth.css
execute:
  echo: false
---

# Initialization

The initial values and variables are loaded from the file stored in variable `identif` in a different structure as before.

```{r packages, warning = F, message = F}
library(tidyverse)
library(igraph)
library(qgraph)
library(visNetwork)
library(scales)
```

```{r init, warning = FALSE, message = FALSE}
root <- rprojroot::find_root(
  rprojroot::is_rstudio_project | rprojroot::is_git_root
)

variable.table <<- read.csv(
  file.path(root, "data", "Variable_Definitions.csv")
)
sector_list <- read.csv(
  file.path(root, 'data', 'sector_list.csv'),
  stringsAsFactors = FALSE
)

source(file.path(root, "functions", "utils.R"))

identif <- file.path(root, "data", "initial_state_2026.xlsx")
initial <- load.new.init(identif)

source(file.path(root, "functions", "run_model_2026.R"))
source(file.path(root, "functions", "MVP_model_2026.R"))
```

```{r filenames}
# -----------------------------
# FILENAMES
# -----------------------------

# -----------------------------
# Directories
# -----------------------------

shock_dir <- file.path(root, "data", 'scenarios')

dir_runs <- file.path(shock_dir, "shock_runs")
dir_tables <- file.path(shock_dir, "shock_tables")
dir_logs <- file.path(shock_dir, "logs")
dir_long <- file.path(shock_dir, "shock_long")
dir_meta <- file.path(shock_dir, "scenario_cache")

dir.create(shock_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(dir_runs, showWarnings = FALSE, recursive = TRUE)
dir.create(dir_tables, showWarnings = FALSE, recursive = TRUE)
dir.create(dir_logs, showWarnings = FALSE, recursive = TRUE)
dir.create(dir_meta, showWarnings = FALSE, recursive = TRUE)

# -----------------------------
# Filenames
# -----------------------------

baseline_filename <- file.path(shock_dir, 'baseline_feb10.RDS')
target_filename <- file.path(shock_dir, 'target_values.xlsx')
scenario_filename <- file.path(shock_dir, 'scenarios.csv')

get_shock_filename <- function(n) {
  file.path(
    dir_runs,
    paste0('shock_', n, '_run.RDS')
  )
}

get_table_filename <- function(n) {
  file.path(
    dir_tables,
    paste0('shock_', n, '_table.csv')
  )
}

get_log_filename <- function(n) {
  file.path(
    dir_logs,
    paste0('shock_', n, '_run.log')
  )
}
```


# Baseline 

```{r simulation, warning = F, message = F}
# baseline <- run.model(
#   initial,
#   mvp.model,
#   log_file = file.path(root, "logs", "baseline_run.log"),
#   log_append = FALSE,
#   print_final_state = FALSE
# )

# saveRDS(baseline, file = file.path(root, 'data', 'baseline_feb10.RDS'))
baseline <- readRDS(
  file = baseline_filename
)

```

## Calibration

Target values are loaded in format long (`target.set`) and wide (`target.table.0`). Units are divided by 10,000 for convenience.

```{r calibration}
#### LOAD TARGET VALUES ####

target.list <- load.target(target_filename)
target.table <- target.list$wide
target.set <- target.list$long
target.vars <- target.set$var.label
target.set$target <- baseline$simulation[
  target.set$var,
  baseline$initial$pars['nPeriods', 'value']
]
target.set$ratio <- round(target.set$target / target.set$value, 2)

compute.target(baseline)
```

## Visualization

```{r variable_selection} 
#| fig-width: 7
#| fig-height: 4

# Variable Selection
selected.list <- list(
  Macroeconomic = c('c', 'va', 'go', 'id', 'gdef', 'cab', 'tb'),
  Social = c('n', 'nf', 'shp', 'shw'),
  Debt.and.Wealth = c('lh', 'lf', 'v', 'k', 'b_s'),
  Ecological = c('x_mat', 'land', 'mat', 'rec', 'emis', 'wa', 'water')
)

plot.selected.by_group(baseline, selected.list)
```

___

# Scenarios to Analyze

| Domain      | Shift              | From | To | Mechanism                  | Conserved Quantity            |
|-------------|--------------------|------|----|----------------------------|-------------------------------|
| Household Demand   | Green Diet Shift         | 7    | 8  | Final demand reallocation  | Total food expenditure        |
| Household Demand   | Green Energy Shift  | 31   | 32 | Final demand reallocation  | Energy service demand         |
| Production  | Wood circularity   | 11   | 12 | Technical coefficient change | Total wood material input   |
| Production  | Pulp circularity   | 13   | 14 | Technical coefficient change | Paper output                |
| Production  | Plastics circularity | 17 | 18 | Technical coefficient change | Plastic product output      |
| Production  | Metal circularity  | 26   | 27 | Technical coefficient change | Metal product output        |
| Production  | Glass circularity  | 21   | 22 | Technical coefficient change | Glass product output        |
| Production  | Cement circularity | 24   | 25 | Technical coefficient change | Cementitious output         |
| Production  | Green Energy Shift  | 31   | 32 | Technical coefficient change | Energy service to production|
| Production  | Construction reuse | 36   | 27 | Technical coefficient change | Construction output         |

```{r scenarios.table}
sc <- read.csv(file = scenario_filename)
scenario_list <- list()
```

___

# Scenario 1: Household Demand - Green Diet Shift

```{r shock.run}
# Time Period Selection
t.shock <- initial$pars['t.shock', 'value'] # shock time period
t0 <- 55 # initial time period for visualization
tf <- initial$pars['nPeriods', 'value'] # final time period for visualization

# Shock Selection and Run
initial_shock <- initial
n_shock <- initial_shock$pars['shock', 'value'] <- 1

scenario <- scenario_list[[n_shock]] <- run.model(
  initial_shock,
  mvp.model,
  log_file = get_log_filename(n_shock),
  log_append = FALSE,
  print_final_state = FALSE
)
saveRDS(
  scenario,
  file = get_shock_filename(n_shock)
)

# scenario <- readRDS(
#   file = get_shock_filename(n_shock)
# )
```

## Visualize Shocked Variables

```{r shock.view}
#| fig-width: 6
#| fig-height: 2

view.shock(
  scenario,
  shock.vars = c(
    'beta-7',
    'beta-8'
  ),
  shock.sectors = c(7, 8),
  shock.title = sc$display_name[n_shock]
)
```

## Visualize Impacts

```{r shock.visualize.scaled}
#| fig-width: 8
#| fig-height: 4

# Load Baseline - Shock Data
df <- shock.long.new(baseline, scenario)

# Plot Selected List Variables (scaled only if gdef is not there)
for (i in 1:length(selected.list)) {
  if ('gdef' %in% selected.list[[i]]) {
    print(view.vars(
      data = df,
      viz.vars = selected.list[[i]],
      var.label = names(selected.list)[i],
      shock.title = sc$display_name[n_shock]
    ))
  } else {
    print(view.scaled.vars(
      data = df,
      viz.vars = selected.list[[i]],
      var.label = names(selected.list)[i],
      shock.title = sc$display_name[n_shock]
    ))
  }
}

# Print Some Scaled Variables Example
# view.scaled.vars(data = shock.long.new(baseline, ple.shock),
#                  viz.vars = c('cab', 'ydw', 'n', 'rva'),
#                  var.label = 'Scaled')

# Print Some Variables Example (not scaled)
# view.vars(data = shock.long.new(baseline, ple.shock),
#           viz.vars = c('cab', 'ydw', 'n', 'rva'),
#           var.label = 'Not Scaled')
```

## Visualize Selected Variables

```{r shock.visualize.selected.1}
#| fig-width: 8
#| fig-height: 4

# Selected Vars to plot
selected <- c('n', 'c', 'shw', 'va', 'cab', 'nf', 'gdef', 'tb', 'go')

p.selected.vars <- filter(df, variable %in% selected) %>%
  ggplot(aes(x = time, y = value, color = area, linetype = scenario)) +
  facet_wrap(~name, scales = 'free') +
  geom_line() +
  geom_vline(
    xintercept = initial$pars['t.shock', 'value'],
    linetype = 'dashed',
    linewidth = .4
  ) +
  labs(
    title = sc$display_name[n_shock],
    subtitle = 'Selected Indicators. Vertical dashed line indicates shock time'
  ) +
  theme(
    plot.title = element_text(face = 'bold'),
    plot.subtitle = element_text(face = 'italic', size = 10),
    legend.title = element_text(face = 'bold')
  )

p.selected.vars
```

```{r summary.table.1}
#| fig-width: 10
#| fig-height: 10

scenario.table <- shock.summary(
  baseline,
  shock.run = scenario,
  t_ = c(t.shock + 1, t.shock + 7, t.shock + 20),
  t_names = c('Immediate', 'Short.Term', 'Long.Term')
)

# kable(diet.table, booktabs = T)
write.csv(scenario.table, get_table_filename(n_shock), row.names = FALSE)

df <- scenario <- read.csv(get_table_filename(n_shock))
source(file.path(root, 'R', 'scenario_analysis_2026.R'))
```

___

# Scenario 2: Household Demand - Green Energy Shift

```{r shock.run.2}
# Time Period Selection
t.shock <- initial$pars['t.shock', 'value'] # shock time period
t0 <- 55 # initial time period for visualization
tf <- initial$pars['nPeriods', 'value'] # final time period for visualization

# Shock Selection and Run
initial_shock <- initial
initial_shock$pars['shock', 'value'] <- 2

# energy.shock <- run.model(
#   initial_shock,
#   mvp.model,
#   log_file = file.path(root, "logs", "shock_2_run.log"),
#   log_append = FALSE,
#   print_final_state = FALSE
# )
# saveRDS(
#   energy.shock,
#   file = file.path(root, 'data', 'shock_2_energy_shock.RDS')
# )

energy.shock <- readRDS(
  file = file.path(root, sc$rds_file[n_shock])
)
```

## Visualize Shocked Variables

```{r shock.view.2}
#| fig-width: 6
#| fig-height: 2

view.shock(
  energy.shock,
  shock.vars = c(
    'beta-31',
    'beta-32'
  ),
  shock.sectors = c(31, 32),
  shock.title = sc$display_name[n_shock]
)
```

## Visualize Impacts

```{r shock.visualize.scaled.2}
#| fig-width: 8
#| fig-height: 4

# Load Baseline - Shock Data
df <- shock.long.new(baseline, energy.shock)

# Plot Selected List Variables (scaled only if gdef is not there)
for (i in 1:length(selected.list)) {
  if ('gdef' %in% selected.list[[i]]) {
    print(view.vars(
      data = df,
      viz.vars = selected.list[[i]],
      var.label = names(selected.list)[i],
      shock.title = sc$display_name[n_shock]
    ))
  } else {
    print(view.scaled.vars(
      data = df,
      viz.vars = selected.list[[i]],
      var.label = names(selected.list)[i],
      shock.title = sc$display_name[n_shock]
    ))
  }
}

# Print Some Scaled Variables Example
# view.scaled.vars(data = shock.long.new(baseline, ple.shock),
#                  viz.vars = c('cab', 'ydw', 'n', 'rva'),
#                  var.label = 'Scaled')

# Print Some Variables Example (not scaled)
# view.vars(data = shock.long.new(baseline, ple.shock),
#           viz.vars = c('cab', 'ydw', 'n', 'rva'),
#           var.label = 'Not Scaled')
```

## Visualize Selected Variables

```{r shock.visualize.selected.2}
#| fig-width: 8
#| fig-height: 4

# Selected Vars to plot
selected <- c('n', 'c', 'shw', 'va', 'cab', 'nf', 'gdef', 'tb', 'go')

p.selected.vars <- filter(df, variable %in% selected) %>%
  ggplot(aes(x = time, y = value, color = area, linetype = scenario)) +
  facet_wrap(~name, scales = 'free') +
  geom_line() +
  geom_vline(
    xintercept = initial$pars['t.shock', 'value'],
    linetype = 'dashed',
    linewidth = .4
  ) +
  labs(
    title = sc$display_name[n_shock],
    subtitle = 'Selected Indicators. Vertical dashed line indicates shock time'
  ) +
  theme(
    plot.title = element_text(face = 'bold'),
    plot.subtitle = element_text(face = 'italic', size = 10),
    legend.title = element_text(face = 'bold')
  )

p.selected.vars
```

## Summary Visualization

```{r summary.table.2}
energy.table <- shock.summary(
  baseline,
  shock.run = energy.shock,
  t_ = c(t.shock + 1, t.shock + 7, t.shock + 20),
  t_names = c('Immediate', 'Short.Term', 'Long.Term')
)

# kable(energy.table, booktabs = T)
write.csv(energy.table, get_table_filename(n_shock), row.names = FALSE)
```

```{r summary_plot_2}
#| fig-width: 10
#| fig-height: 10

df <- energy.table <- read.csv(get_table_filename(n_shock))
source(file.path(root, 'R', 'scenario_analysis_2026.R'))
```

___

# Scenario 3: Production - Wood Shift

```{r shock.run.3}
# Time Period Selection
t.shock <- initial$pars['t.shock', 'value'] # shock time period
t0 <- 55 # initial time period for visualization
tf <- initial$pars['nPeriods', 'value'] # final time period for visualization

# Shock Selection and Run
initial_shock <- initial
initial_shock$pars['shock', 'value'] <- 3

# wood.shock <- run.model(
#   initial_shock,
#   mvp.model,
#   log_file = file.path(root, "logs", "shock_3_run.log"),
#   log_append = FALSE,
#   print_final_state = FALSE
# )
# saveRDS(
#   wood.shock,
#   file = file.path(root, 'data', 'shock_runs', 'shock_3_run.RDS')
# )

wood.shock <- readRDS(
  file = file.path(root, sc$rds_file[n_shock])
)
```

## Visualize Shocked Variables

```{r shock.view.3}
#| fig-width: 6
#| fig-height: 2

view.shock(
  wood.shock,
  shock.vars = c(
    'beta-31',
    'beta-32'
  ),
  shock.sectors = c(31, 32),
  shock.title = sc$display_name[n_shock]
)
```

## Visualize Impacts

```{r shock.visualize.scaled.3}
#| fig-width: 8
#| fig-height: 4

# Load Baseline - Shock Data
df <- shock.long.new(baseline, energy.shock)

# Plot Selected List Variables (scaled only if gdef is not there)
for (i in 1:length(selected.list)) {
  if ('gdef' %in% selected.list[[i]]) {
    print(view.vars(
      data = df,
      viz.vars = selected.list[[i]],
      var.label = names(selected.list)[i],
      shock.title = sc$display_name[n_shock]
    ))
  } else {
    print(view.scaled.vars(
      data = df,
      viz.vars = selected.list[[i]],
      var.label = names(selected.list)[i],
      shock.title = sc$display_name[n_shock]
    ))
  }
}

# Print Some Scaled Variables Example
# view.scaled.vars(data = shock.long.new(baseline, ple.shock),
#                  viz.vars = c('cab', 'ydw', 'n', 'rva'),
#                  var.label = 'Scaled')

# Print Some Variables Example (not scaled)
# view.vars(data = shock.long.new(baseline, ple.shock),
#           viz.vars = c('cab', 'ydw', 'n', 'rva'),
#           var.label = 'Not Scaled')
```

## Visualize Selected Variables

```{r shock.visualize.selected.3}
#| fig-width: 8
#| fig-height: 4

# Selected Vars to plot
selected <- c('n', 'c', 'shw', 'va', 'cab', 'nf', 'gdef', 'tb', 'go')

p.selected.vars <- filter(df, variable %in% selected) %>%
  ggplot(aes(x = time, y = value, color = area, linetype = scenario)) +
  facet_wrap(~name, scales = 'free') +
  geom_line() +
  geom_vline(
    xintercept = initial$pars['t.shock', 'value'],
    linetype = 'dashed',
    linewidth = .4
  ) +
  labs(
    title = sc$display_name[n_shock],
    subtitle = 'Selected Indicators. Vertical dashed line indicates shock time'
  ) +
  theme(
    plot.title = element_text(face = 'bold'),
    plot.subtitle = element_text(face = 'italic', size = 10),
    legend.title = element_text(face = 'bold')
  )

p.selected.vars
```

## Summary Visualization

```{r summary.table.3}
shock_3_table <- shock.summary(
  baseline,
  shock.run = wood.shock,
  t_ = c(t.shock + 1, t.shock + 7, t.shock + 20),
  t_names = c('Immediate', 'Short.Term', 'Long.Term')
)

# kable(shock_3_table, booktabs = T)
write.csv(shock_3_table, get_table_filename(n_shock), row.names = FALSE)
```

```{r summary_plot_3}
#| fig-width: 10
#| fig-height: 10

df <- shock_3_table <- read.csv(get_table_filename(n_shock))
source(file.path(root, 'R', 'scenario_analysis_2026.R'))
```

___

# Scenario Comparison

```{r comparison} 
#| fig-width: 10
#| fig-height: 21
#| fig-cap: "Comparison of Diet and Energy Transition Scenarios"

selected.list <- list(
  Macroeconomic = c('c', 'va', 'go', 'id', 'gdef', 'cab', 'tb'),
  Social = c('n', 'nf', 'shp', 'shw'),
  Debt.and.Wealth = c('lh', 'lf', 'v', 'k', 'b_s'),
  Ecological = c('x_mat', 'land', 'mat', 'rec', 'emis', 'wa', 'water')
)

lookup <- stack(selected.list) |>
  dplyr::rename(
    Variable = values,
    Name = ind
  )

scenario.tables <- list(
  diet = diet.table,
  energy = energy.table,
  wood = shock_3_table
) %>%
  bind_rows(.id = "scenario") %>%
  mutate(
    scenario = factor(
      scenario,
      levels = c("diet", "energy", 'wood'),
      labels = c(
        'Final Demand\nHousehold Diet Shift',
        'Final Demand\nHousehold Consumption Energy Shift',
        'Intermediate Demand\nWood Shift'
      )
    )
  )

df <- scenario.tables |>
  dplyr::select(-Name) |> # drop old category column
  dplyr::left_join(lookup, by = "Variable") %>%
  filter(!is.na(Name))

source(file.path(root, "R", "scenario_comparison_2026.R"))
p.comparison
```

___

# Key Structural Findings from Scenario Comparison

The scenario analysis reveals a clear macro-ecological success combined with persistent social and distributional tensions. The green transition improves environmental indicators and external balances but generates labour market contraction and rising inequality, particularly over time.

## Macroeconomic Stabilization

Across both Core and Periphery, the green transition produces systematic macroeconomic improvements, especially beyond the immediate adjustment phase.

- Current account balance and trade balance improve in all time horizons  
- Government deficit narrows significantly in the short and long term  
- Value added and gross output recover after short-term adjustment costs  
- The Core captures stronger and earlier gains than the Periphery  

**Time dynamics:**  
Short-term disruption → Medium-term stabilization → Long-term consolidation  

These patterns are consistent with a structural reallocation process rather than a prolonged recessionary adjustment.

---

## Distributional Shift Toward Capital

The transition generates a persistent and intensifying redistribution from labour to capital.

| Indicator | Immediate | Short Term | Long Term | Interpretation |
|----------|-----------|------------|-----------|----------------|
| Wage share of income | ↓ | ↓↓ | ↓↓ | Labour loses income share |
| Profit share of income | ↑ | ↑↑ | ↑↑ | Capital captures transition rents |
| Worker disposable income | ↓ | ↓ | ↓ | Real labour income compression |
| Capitalist disposable income | ↑ | ↑↑ | ↑↑ | Income polarization widens |

The shift is not temporary; it strengthens over time, indicating a structural, capital-biased transition under existing institutional settings.

---

## Labour Market Contraction

Employment effects are consistently negative, with no clear recovery in the long term.

| Variable | Core | Periphery | Dynamics |
|----------|------|-----------|----------|
| Total employment | ↓ | slight ↓ | Persistent decline |
| Female employment | ↓↓ | slight ↓ | Gendered impact |
| Real consumption | ↓ | flat/slight ↓ | Demand compression |

The results suggest a labour-saving restructuring rather than a short-term displacement followed by reabsorption. The stronger fall in female employment points to sectoral reallocation away from feminized sectors.

---

## Core–Periphery Asymmetry

The macroeconomic benefits of the transition are unevenly distributed:

- The Core experiences stronger improvements in trade balances and current account positions  
- Financial variables and wealth indicators suggest capital re-accumulation in the Core  
- The Periphery benefits less from value-added gains  

This pattern indicates upstream green upgrading concentrated in core economies, with limited peripheral value capture.

---

## Strong and Immediate Ecological Gains

Environmental improvements are large, immediate, and sustained across all horizons.

| Indicator | Pattern |
|----------|---------|
| CO₂ emissions | Large decline |
| Material extraction | Decline |
| Land use | Decline |
| Waste | Decline |
| Water use | Decline |
| Recycling | Increase |

Unlike the social variables, ecological gains are front-loaded rather than delayed.

---

## Time Horizon Summary

| Dimension | Immediate | Short Term | Long Term |
|-----------|-----------|------------|-----------|
| Ecology | Large gains | Maintained | Maintained |
| Macroeconomic stability | Mixed | Improves | Consolidates |
| Employment | Falls | Falls further | No recovery |
| Inequality | Rises | Rises more | Structural shift |
| Core–Periphery gap | Limited | Widening | Entrenched |

---

## Overall Interpretation

The modeled green transition is:

**Ecologically effective**  
**Macroeconomically stabilizing**  
**Socially regressive under current institutions**  

Without complementary policies such as:

- Wage protection and collective bargaining support  
- Employment guarantees or green job programs  
- Labour-absorbing industrial policy  
- Gender-sensitive transition measures  

the transition path risks becoming politically destabilizing despite ecological success.

---

## Executive Takeaway

- The green transition delivers rapid ecological improvements and medium-term macroeconomic stabilization, particularly in external balances and fiscal positions. 
- However, these gains are accompanied by persistent labour displacement, declining wage shares, rising profit shares, and falling worker incomes, with especially strong negative effects on female employment. 
- Inequality widens over time, and the Core captures a disproportionate share of the macroeconomic benefits relative to the Periphery.
- Without strong redistributive and labour-market institutions, the transition appears environmentally effective but socially and politically fragile.

