---
title: "ECO-MRIO-SFC Model: 54-Sector Version Documentation"
author: "Oriol Vallès Codina"
date: "`r Sys.Date()`"
output:
  # word_document: default
  # keep_tex: true

  # pdf_document:
  #   toc: true
  #   toc_depth: 3
  #   number_sections: true
  html_document:
    toc: TRUE
    toc_float: TRUE
format:
  html:
    self-contained:
      true
---

# Initialization

The initial values and variables are loaded from the file stored in variable `identif` in a different structure as before.

```{r init, warning = FALSE, message = FALSE}
# Clear Environment
rm(list = ls())

# Load Auxiliary Functions
directory <- '' # '' for the markdown file, 'flexible_/' for R
variable.table <<- read.csv(paste0(directory, 'data/Variable_Definitions.csv'))
source(paste0(directory, 'functions/auxiliary_flexible_aug12.R')) # load.new.init, auxiliary labeling functions, plotting function

# Load Initial Values
identif <- paste0(directory, 'data/perfect_init_aug9_withBmatrix.xlsx')
initial <- load.new.init(identif)

# Load Models
source(paste0(directory, "functions/run_new_model_aug12.R")) # function run.new.model(initial.values, model.equations)
source(paste0(directory, "functions/MVP_model_aug12.R")) # model equations

baseline <- run.new.model(initial, mvp.model)
saveRDS(baseline, file = 'data/baseline_aug13.RDS')
# baseline <- readRDS(file = 'data/baseline_aug9.RDS')
```

# Table of Industries

```{r industry.table}
kable(data.frame(number = 1 : length(initial$sectors), industry = initial$sectors), 
      booktabs = T)
```


# MRIO $A$ Matrix Visualization

Importantly, the column names of the $A$ matrix contained in `identif` determine the country and industry labels in the whole model, following the grammar `country_label.industry_label`:

```{r network}
library(igraph)
library(qgraph)
go <- rowSums(initial$A.matrix)
g <- graph_from_adjacency_matrix(as.matrix(initial$A.matrix), 
                                 mode = 'directed', weighted = TRUE,diag = F, add.colnames = T)

# Graph Visualization
l <- qgraph.layout.fruchtermanreingold(get.edgelist(g, names=FALSE), vcount=vcount(g))
V(g)$size <- 7.5 * go ^ 0.75
V(g)$frame.color <- NA
# V(g)$name <- apply(expand.grid(1 : unique(initial$K), initial$countries)[c(2,1)], 1, function(x) paste0(x, collapse = ""))
V(g)$name <- rep(1 : unique(initial$K), times = 2)
V(g)$color <- rep(scales :: hue_pal()(4)[c(1, 3)], each = unique(initial$K))
E(g)$color <- alpha('black', E(g)$weight / max(E(g)$weight))

par(mar = c(0,0,0,0))
plot(g, edge.arrow.size = 0)
# title("MRIO (EU, red; RoW, emerald)")
```


# Calibration

Target values are loaded in format long (`target.set`) and wide (`target.table.0`). Units are divided by 10,000 for convenience.

```{r target}
#### LOAD TARGET VALUES ####
target.list <- load.target(paste0(directory, 'data/target_values.xlsx'))
target.table <- target.list$wide
target.set <- target.list$long
target.vars <- target.set$var.label

compute.target(baseline)
```

# Visualization

```{r baseline.viz}
plot.vars(baseline, c(target.vars, 'ineq', 'debt_gdp'))
```

# Shocks

```{r selected.variables}
selected.list <- list(All = c('n', 'c', 'ineq', 'va', 'cab', 'nf', 'gdef', 'tb', 'go'),
                      Macroeconomic = c('c', 'rva', 'go', 'id'),
                      External = c('cab', 'gdef', 'tb'),
                      Social = c('n', 'nf', 'ineq'),
                      Prices = c('pa', 'pid', 'pg', 'pim'),
                      Employment = c('n', 'nf'),
                      Inequality = c('ydw', 'ydc', 'ineq'),
                      Debt.and.Wealth = c('lh', 'lf', 'v', 'k', 'b_s'))

selected <- c('n', 'c', 'ineq', 'va', 'cab', 'nf', 'gdef', 'tb', 'go')
macro.vars<-c('c', 'rva', 'go','id')
external.vars<-c('cab','gdef','tb')
social.vars <- c('n', 'nf', 'ineq')
prices.vars<- c('pa','pid','pg','pim')
employment.vars<-c('n', 'nf')
inequality.vars<-c('ydw','ydc','ineq')
debt.wealth.vars<-c('lh','lf','v','k','bs')
```


# Shock Example 7: Product Lifetime Extension

## Run Shock

This is a simple visualization of shock 7. Please note the new grammar to change initial values.

```{r shock.run}
# Time Period Selection
t.shock <- initial$pars['t.shock', 'value'] # shock time period
t0 <- 50 # initial time period for visualization
tf <- initial$pars['nPeriods', 'value'] # final time period for visualization

# Shock Selection and Run
initial$pars['shock', 'value'] <- 7 # shock 7
ple.shock <- run.new.model(initial, mvp.model)
# saveRDS(ple.shock, file = 'data/pleshock_aug12.RDS')
ple.shock <- readRDS(file = 'data/pleshock_aug12.RDS')

shock.title <- 'Shock: Product Lifetime Extension'
df <- shock.long.new(baseline, ple.shock)
```

## Visualize Shocked Variables

```{r shock.view}
view.shock(ple.shock,
           shock.vars = c('id0', 'iota_g-1', 'iota_g-46', 'zeta_dc-28', 'beta-28', 'beta-29', 'beta-40'),
           shock.sectors = c(NA, NA, 46, 28, 28, 29, 40))
```

## Visualize Scaled Variables

```{r shock.visualize.selected}
# Selected Vars to plot
view.scaled.vars(data = df,
          viz.vars = selected,
          var.label = '')

# Macroeconomic Vars to plot
view.scaled.vars(data = df,
          viz.vars = macro.vars,
          var.label = 'Macroeconomic')

# External Vars to plot
view.scaled.vars(data = df,
          viz.vars = external.vars,
          var.label = 'External')

# Social Vars to plot
view.scaled.vars(data = df,
          viz.vars = social.vars,
          var.label = 'Social')

# Price Vars
view.scaled.vars(data = df,
          viz.vars = c('pa', 'pid', 'pg', 'pim'),
          var.label = 'Price Indices')

# Employment
view.scaled.vars(data = df,
          viz.vars = c('n', 'nf'),
          var.label = 'Employment')

# Inequality
view.scaled.vars(data = df,
          viz.vars = inequality.vars,
          var.label = 'Inequality')

# Debt and Wealth
view.scaled.vars(data = df,
          viz.vars = debt.wealth.vars,
          var.label = 'Financial Stocks')
```

## Scaled Visualization

This is a scaled visualization of shock 7.

```{r scaled.viz.shock}
# Variable Selection
selected.vars <- list(Macroeconomic = c('c', 'va', 'go','id'),
                      External = c('cab', 'gdef', 'tb', 'nimp'),
                      Social = c('n', 'nf', 'ineq', 'pa'))

df <- shock.long.new(baseline, ple.shock)
p.ple.shock <- list()
for (i in length(selected.vars)) p.ple.shock[[i]] <- view.scaled.vars(data = df,
                                                            viz.vars = selected.vars[[i]],
                                                            var.label = names(selected.vars)[i])

# Print Plots
p.ple.shock[[2]]
```

# Shock Example 4: Changing IO Coefficients

## Run Shock

This is a simple visualization of shock 4. Please note the new grammar to change initial values.

```{r shock.run.2}
# Shock Selection and Run
init.ce <- load.new.init('data/perfect_init_aug11_withBmatrix_CE_inputs_higher_costs.xlsx')

# Time Period Selection
t.shock <- init.ce$pars['t.shock', 'value'] # shock time period
t0 <- 50 # initial time period for visualization
tf <- init.ce$pars['nPeriods', 'value'] # final time period for visualization

init.ce$pars['shock', 'value'] <- 4

# ce.shock <- run.new.model(init.ce, mvp.model)
# saveRDS(ce.shock, file = 'data/CE_shock_aug12.RDS')
ce.shock <- readRDS(file = 'data/CE_shock_aug12.RDS')

shock.title <- 'Shock: Change in IO Coefficients'
df <- shock.long.new(baseline, ce.shock)
```

## Visualize Selected Variables

```{r shock.visualize.selected.2}
# Selected Vars to plot
selected <- c('n', 'c', 'ineq', 'va', 'cab', 'nf', 'gdef', 'tb', 'go')

p.selected.vars <- filter(df, variable %in% selected) %>% 
    ggplot(aes(x = time, y = value, color = area, linetype = scenario)) + facet_wrap(~ name, scales = 'free') + geom_line() + 
    geom_vline(xintercept = initial$pars['t.shock', 'value'], linetype = 'dashed', linewidth = .4) +
    labs(title = shock.title, subtitle = 'Selected Indicators. Vertical dashed line indicates shock time')

p.selected.vars
```

## Visualize Scaled Variables

```{r shock.visualize.selected.3}
# Macroeconomic Vars to plot
view.scaled.vars(data = df,
          viz.vars = macro.vars,
          var.label = 'Macroeconomic')

# External Vars to plot (not scaled)
view.vars(data = df,
          viz.vars = external.vars,
          var.label = 'External')

# Social Vars to plot
view.scaled.vars(data = df,
          viz.vars = social.vars,
          var.label = 'Social')

# Price Vars
view.scaled.vars(data = df,
          viz.vars = c('pa', 'pid', 'pg', 'pim'),
          var.label = 'Price Indices')

# Employment
view.scaled.vars(data = df,
          viz.vars = c('n', 'nf'),
          var.label = 'Employment')

# Inequality
view.scaled.vars(data = df,
          viz.vars = inequality.vars,
          var.label = 'Inequality')

# Debt and Wealth
view.scaled.vars(data = df,
          viz.vars = debt.wealth.vars,
          var.label = 'Financial Stocks')
```

# Summary Tables for the Scenarios

```{r}
# Variable Selection
selected.list <- list(All = c('n', 'c', 'ineq', 'va', 'cab', 'nf', 'gdef', 'tb', 'go'),
                      Macroeconomic = c('c', 'rva', 'go', 'id'),
                      External = c('cab', 'gdef', 'tb'),
                      Social = c('n', 'nf', 'ineq'),
                      Prices = c('pa', 'pid', 'pg', 'pim'),
                      Employment = c('n', 'nf'),
                      Inequality = c('ydw', 'ydc', 'ineq'),
                      Debt.and.Wealth = c('lh', 'lf', 'v', 'k', 'b_s'))

# Tables
ple.table <- shock.summary(baseline, 
                           shock.run = ple.shock, 
                           t_ = c(t.shock + 1, t.shock + 7, t.shock + 20),
                           t_names = c('Immediate', 'Short.Term', 'Long.Term'))

kable(ple.table, booktabs = T)

ce.table <- shock.summary(baseline, 
                          shock.run = ce.shock, 
                          t_ = c(t.shock + 1, t.shock + 7, t.shock + 20),
                          t_names = c('Immediate', 'Short.Term', 'Long.Term'))

kable(ce.table, booktabs = T)
```

