---
title: "Effective climate action can only be global in scope"
subtitle: "Insights from a Core-Periphery SOCIO-ECO-MRIO-SFC Model"
author:
  - name: "Oriol Vallès Codina"
    affiliation: "Johns Hopkins University, Net Zero Industrial Policy Lab, Department of Political Science"
    email: 
      - "oriolvallescodina@gmail.com"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc_float: true
    self-contained: true
    page-layout: full
    css: fullwidth.css
execute:
  echo: false
---

# Initialization

The initial values and variables are loaded from the file stored in variable `identif` in a different structure as before.

```{r packages, warning = F, message = F}
library(tidyverse)
library(igraph)
library(qgraph)
library(visNetwork)
library(scales)
```

```{r init, warning = FALSE, message = FALSE}

# Load Auxiliary Functions
directory <- '' # '' for the markdown file, 'flexible_/' for R
variable.table <<- read.csv('data/Variable_Definitions.csv')
# source('functions/auxiliary_flexible_aug12.R') # load.new.init, auxiliary labeling functions, plotting
source('functions/auxiliary_flexible_nov25.R')


# Load Initial Values
identif <- 'data/initial_state_2026.xlsx' # is the same of 'data/perfect_init_aug14_withBmatrix_PLE_CE.xlsx'
initial <- load.new.init(identif)

# Load Models
source("functions/run_new_model_feb7.R") # function run.new.model(initial.values, model.equations)
source("functions/MVP_model_aug14.R") # model equations

# baseline <- run.new.model(initial, mvp.model)
# saveRDS(baseline, file = 'data/baseline_nov25.RDS')
baseline <- readRDS(file = 'data/baseline_nov25.RDS')
```


# Network Analysis of the MRIO $A$ Matrix

## Static Visualization

Importantly, the column names of the $A$ matrix contained in `identif` determine the country and industry labels in the whole model, following the grammar `country_label.industry_label`:

```{r network_setup}
sector_list <- read.csv('data/sector_list.csv', stringsAsFactors = FALSE)
# convert the literal "<br>" sequences to actual newline characters
sector_list$name <- gsub("<br>", " ", sector_list$label, fixed = TRUE)
sector_list$label <- gsub("<br>", "\n", sector_list$label, fixed = TRUE)

ntwrk <- as.matrix(initial$A.matrix)
n <- nrow(ntwrk)
n_sectors <- nrow(sector_list)
n_regions <- n / n_sectors

regions <- c("EU", "RoW")
region <- rep(regions, each = n_sectors)
sector_names <- apply(
  expand.grid(sector_list$name, paste0('(', regions, ')')),
  1,
  paste,
  collapse = ' '
)


pal <- scales::hue_pal()(4)
region_colors <- c(
  EU = pal[1], # same red as igraph
  RoW = pal[3] # same green/emerald as igraph
)
```


```{r network_static}
#| fig-width: 10
#| fig-height: 10

go <- rowSums(initial$A.matrix)
df <- read.csv('data/impact_coefficients.csv')
go <- setNames(df$Gross.Output, paste0(df$region, '.', df$sector))

g <- graph_from_adjacency_matrix(
  ntwrk,
  mode = 'directed',
  weighted = TRUE,
  diag = F,
  add.colnames = T
)

# Graph Visualization
l <- qgraph.layout.fruchtermanreingold(
  get.edgelist(g, names = FALSE),
  vcount = vcount(g)
)
V(g)$size <- 1e-3 * go^0.6
V(g)$frame.color <- NA
# V(g)$name <- apply(expand.grid(1 : unique(initial$K), initial$countries)[c(2,1)], 1, function(x) paste0(x, collapse = ""))
V(g)$name <- rep(1:n_sectors, times = 2)
V(g)$color <- rep(region_colors, each = n_sectors)
E(g)$color <- alpha('black', E(g)$weight / max(E(g)$weight))

par(mar = c(0, 0, 0, 0))
plot(g, edge.arrow.size = 0)
# title("MRIO (EU, red; RoW, emerald)")

# centralities <- eigen_centrality(
#   g,
#   directed = TRUE,
#   weights = E(g)$weight
# )$vector
# centralities <- setNames(centralities, sector_names)
```

## Dynamic Visualization

```{r visnetwork}
#| fig-width: 10
#| fig-height: 10
#|
diag(ntwrk) <- 0

labels <- rep(sector_list$label, length.out = n)

# ---- nodes ----
nodes <- data.frame(
  id = seq_len(n),
  label = labels,
  value = go,
  group = region,
  color = region_colors[region],
  stringsAsFactors = FALSE
)

# ---- edge threshold: top 10% (given) ----
thr <- 0.001

idx <- which(ntwrk > thr, arr.ind = TRUE)
stopifnot(nrow(idx) > 0)

w <- ntwrk[idx]
w_max <- max(w)

# rescale above threshold
w_scaled <- (w - thr) / (w_max - thr)
w_scaled[w_scaled < 0] <- 0

# nonlinear alpha (visibility kicks in above threshold)
gamma_alpha <- 1.05
alpha_vals <- w_scaled^gamma_alpha

edges <- data.frame(
  from = idx[, "row"],
  to = idx[, "col"],
  value = w,
  width = 0.5 + .5 * w_scaled,
  color = alpha("steelblue", alpha_vals),
  stringsAsFactors = FALSE
)

# ---- visNetwork ----
visNetwork(nodes, edges, height = "700px", width = "100%") %>%
  visNodes(
    shape = "dot",
    scaling = list(min = 2, max = 20),
    font = list(size = 7)
  ) %>%
  visEdges(
    arrows = list(
      to = list(enabled = FALSE),
      from = list(enabled = FALSE)
    ),
    smooth = FALSE
  ) %>%
  visOptions(
    highlightNearest = TRUE,
    nodesIdSelection = TRUE
  ) %>%
  visPhysics(enabled = FALSE)

```

## Top Nodes by Size (Top Industries by Gross Output)

```{r top_nodes_by_size}
#| fig-width: 10
#| fig-height: 8

node_size <- go

top_n <- 20
ix <- order(node_size, decreasing = TRUE)[1:top_n]


top_df <- node_size[ix] |>
  as.data.frame() |>
  rownames_to_column(var = "node") %>%
  mutate(
    region = sub("\\..*$", "", node),
    sector = sub("^[^.]*\\.", "", node),
    name = sector_names[ix]
  )

names(top_df)[2] <- "size"

p <- ggplot(top_df, aes(y = reorder(name, size), x = size, fill = region)) +
  geom_bar(stat = "identity") +
  labs(
    y = '',
    x = 'value (million $)',
    title = 'Top Industries by Gross Output'
  ) +
  theme(plot.title = element_text(face = 'bold', size = 14))

p

png('figures/top20nodes.png')
p
dev.off()

top_nodes <- top_df
```


## Top Edges

```{r}
#| fig-width: 9
#| fig-height: 7.5

edges <- edges |>
  mutate(
    from.name = sector_names[from],
    to.name = sector_names[to]
  )

# Infer region from the sector name prefix
edges <- edges |>
  mutate(
    from.region = region[from],
    to.region = region[to],
    flow_type = paste(from.region, to.region, sep = "–") # "EU–EU", "EU–RoW", etc.
  )

top20 <- edges |>
  slice_max(value, n = 20) |>
  mutate(
    link = paste0(from.name, " → ", to.name)
  )


p <- ggplot(
  top20,
  aes(
    x = value,
    y = reorder(link, value),
    fill = flow_type
  )
) +
  geom_col() +
  labs(
    x = "Weight",
    y = "",
    fill = "Region flow",
    title = "Top 20 Links by IO Coefficients",
    subtitle = 'Colored by region-region flows'
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    axis.text.y = element_text(size = 10)
  )

p

png('figures/top20links.png', width = 600, height = 480)
p
dev.off()

# No production links from EU to ROW in the top20
print('top20 links by region-region flow')
table(top20$from.region, top20$to.region)

print('percentage out of all links by region-region flow')
round(100 * table(edges$from.region, edges$to.region) / nrow(edges), digits = 2)

top_links <- top20
```

> Out of the top 20 links, 14 are RoW-RoW production links, 5 are EU-EU production links, and only 1 is a cross-border link from RoW to the EU, the second largest IO coefficient: Fossil-Fuel Extraction (RoW) to Fossil-Fuel Processing (EU). 

> There are 3 times more production links from RoW to the EU than from the EU to RoW.

```{r top20links_RoW_EU}
#| fig-width: 9
#| fig-height: 7.5

top20 <- edges |>
  filter(from.region == "RoW" & to.region == "EU") |>
  slice_max(value, n = 20) |>
  mutate(
    link = paste0(from.name, " → ", to.name)
  )

ggplot(
  top20,
  aes(
    x = value,
    y = reorder(link, value),
    fill = flow_type
  )
) +
  geom_col(show.legend = FALSE) +
  labs(
    x = "Weight",
    y = "",
    fill = "Region flow",
    title = "Top 20 RoW-EU Links by Weight"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    axis.text.y = element_text(size = 10)
  )

top_crosslinks <- top20
```

## Top Nodes by Centrality

```{r top_nodes_by_centrality, eval = F}
#| fig-width: 10
#| fig-height: 10

ec <- eigen_centrality(
  g,
  directed = TRUE,
  weights = E(g)$weight
)$vector

top_n <- length(ix) # reuse same N if you want consistency
ix_ec <- order(ec, decreasing = TRUE)[1:top_n]

top_ec_df <- ec[ix_ec] |>
  as.data.frame() |>
  rownames_to_column(var = "node") %>%
  mutate(
    region = sub("\\..*$", "", node),
    sector = sub("^[^.]*\\.", "", node),
    label = sector_names[ix_ec]
  )

names(top_ec_df)[2] <- "eigenvector"

ggplot(
  top_ec_df,
  aes(
    y = reorder(label, eigenvector),
    x = eigenvector,
    fill = region
  )
) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    x = "",
    y = "",
    title = "Top Nodes by Eigenvector Centrality"
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 14)
  )

```


## Intensities

### Carbon Intensity

```{r intensities}
#| fig-width: 10
#| fig-height: 10

intensity_names <- c(
  'Share in Household Consumption',
  # 'Final Demand in Real Terms',
  'Carbon Intensity',
  'Share of Imports',
  'Share of Investment',
  'Matter Intensity of Output'
  # 'Female Share of Employment',
  #  'Productivity'
)

intensity_names <- c(
  'Carbon Intensity',
  'Matter Intensity of Output'
)


# First, determine the order based on sector numbers

intensities <- initial$variables$industry |>
  filter(name %in% intensity_names) |>
  mutate(
    region = recode(area, Z1 = "EU", Z2 = "RoW")
  ) |>
  pivot_longer(
    cols = matches("^\\d+$"),
    names_to = "sector",
    values_to = "value"
  ) %>%
  mutate(
    sector = as.integer(sector),
    industry = paste0(
      gsub('\\n', ' ', sector_list$label[sector])
    )
  )


ordered_industries <- sector_list$label[sort(unique(intensities$sector))]
ordered_industries <- gsub('\\n', ' ', ordered_industries)

intensities <- intensities |>
  mutate(,
    industry = factor(industry, levels = rev(ordered_industries))
  )

# Then your plot will maintain this order
p <- intensities %>%
  filter(name == 'Carbon Intensity') %>%
  ggplot(aes(x = value, y = industry, fill = region)) +
  geom_bar(stat = 'identity', position = 'dodge', show.legend = FALSE) +
  labs(
    title = 'Carbon Intensity by Industry and Region (Log Scale)',
    y = '',
    x = 'Intensity of Emissions (ton GHG / million $)'
  ) +
  scale_x_log10() +
  theme(plot.title = element_text(face = "bold", size = 14)) +
  facet_grid(~region)

p
png('figures/carbonintensity.png', width = 600, height = 480)
p
dev.off()

p <- intensities %>%
  filter(name == 'Matter Intensity of Output') %>%
  ggplot(aes(x = 1000 * value, y = industry, fill = region)) +
  geom_bar(stat = 'identity', position = 'dodge', , show.legend = FALSE) +
  labs(
    title = 'Matter Intensity of Output (Log Scale)',
    y = '',
    x = 'Domestic Material Extraction Intensity of Output (ton / million $)'
  ) +
  theme(plot.title = element_text(face = "bold", size = 14)) +
  scale_x_log10() +
  facet_grid(~region)

p
png('figures/matterintensity.png', width = 600, height = 480)
p
dev.off()

```

# Intensity-Output Plots

```{r}
#| fig-width: 8.5
#| fig-height: 7

epsilon <- intensities %>%
  filter(name == 'Carbon Intensity') %>%
  mutate(name = paste0(industry, ' (', region, ')'))

top_nodes %>%
  left_join(epsilon %>% select(-region), by = "name") %>%
  ggplot(aes(x = size, y = value, fill = region, size = size)) +
  geom_point(shape = 21, alpha = 0.8) +
  ggrepel::geom_text_repel(
    aes(label = industry),
    size = 3,
    max.overlaps = Inf,
    box.padding = 0.3,
    point.padding = 0.2,
    show.legend = FALSE
  ) +
  labs(
    x = "Gross Output (million $)",
    y = "Carbon Intensity (ton GHG / million $)",
    title = "Top Nodes by Carbon Intensity and Gross Output",
    size = 'Gross Output'
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 14)
  )
```

```{r}
#| fig-width: 10
#| fig-height: 8.5

top_links %>%
  left_join(
    epsilon %>% rename(from_ci = value),
    by = c("from.name" = "name")
  ) %>%
  left_join(
    epsilon %>% rename(to_ci = value),
    by = c("to.name" = "name")
  ) %>%
  ggplot(aes(x = from_ci, y = to_ci, fill = flow_type, size = value)) +
  geom_point(shape = 21, alpha = 0.8) +
  ggrepel::geom_text_repel(
    aes(label = paste0(industry.x, ' → ', industry.y)),
    size = 3,
    max.overlaps = Inf,
    box.padding = 0.3,
    point.padding = 0.2,
    show.legend = FALSE
  ) +
  labs(
    x = "Carbon Intensity of Input Sector (ton GHG / million $)",
    y = "Carbon Intensity of Output Sector (ton GHG / million $)",
    title = "Top Links by Carbon Intensity",
    size = 'Input-Output Coefficient'
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 14)
  ) +
  facet_grid(from.region ~ to.region, scales = 'free')
```


```{r}
#| fig-width: 8.5
#| fig-height: 7

top_crosslinks %>%
  left_join(
    epsilon %>% rename(from_ci = value),
    by = c("from.name" = "name")
  ) %>%
  left_join(
    epsilon %>% rename(to_ci = value),
    by = c("to.name" = "name")
  ) %>%
  ggplot(aes(x = from_ci, y = to_ci, fill = flow_type, size = value)) +
  geom_point(shape = 21, alpha = 0.8) +
  ggrepel::geom_text_repel(
    aes(label = paste0(industry.x, ' → ', industry.y)),
    size = 3,
    max.overlaps = Inf,
    box.padding = 0.3,
    point.padding = 0.2,
    show.legend = FALSE
  ) +
  labs(
    x = "Carbon Intensity of Input Sector (ton GHG / million $)",
    y = "Carbon Intensity of Output Sector (ton GHG / million $)",
    title = "Top Links by Carbon Intensity",
    size = 'Gross Output'
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 14)
  )
```

# Calibration

Target values are loaded in format long (`target.set`) and wide (`target.table.0`). Units are divided by 10,000 for convenience.

```{r target}
#### LOAD TARGET VALUES ####

target.list <- load.target('data/target_values.xlsx')
target.table <- target.list$wide
target.set <- target.list$long
target.vars <- target.set$var.label
target.set$target <- baseline$simulation[
  target.set$var,
  baseline$initial$pars['nPeriods', 'value']
]
target.set$ratio <- round(target.set$target / target.set$value, 2)

compute.target(baseline)
```

# Shock Example 7: Product Lifetime Extension

## Run Shock

This is a simple visualization of shock 7. Please note the new grammar to change initial values.

```{r shock.run}
# Time Period Selection
t.shock <- initial$pars['t.shock', 'value'] # shock time period
t0 <- 55 # initial time period for visualization
tf <- initial$pars['nPeriods', 'value'] # final time period for visualization

# Shock Selection and Run
initial$pars['shock', 'value'] <- 7 # shock 7
# ple.shock <- run.new.model(initial, mvp.model)
# saveRDS(ple.shock, file = 'data/pleshock_aug14.RDS')
ple.shock <- readRDS(file = 'data/pleshock_aug14.RDS')

shock.title <- 'Scenario: Product Lifetime Extension'
```

## Visualize Shocked Variables

```{r shock.view}
view.shock(
  ple.shock,
  shock.vars = c(
    'delta',
    'id0',
    'iota_g-1',
    'iota_g-46',
    'zeta_dc-28',
    'beta-28',
    'beta-29',
    'beta-41'
  ),
  shock.sectors = c(NA, NA, NA, 46, 28, 28, 29, 41)
)
```

## Visualize Impacts

```{r shock.visualize.scaled}
# Variable Selection
selected.list <- list(
  Economic = c('n', 'c', 'ineq', 'va', 'cab', 'nf', 'gdef', 'tb', 'go'),
  Macroeconomic = c('c', 'rva', 'go', 'id'),
  External = c('cab', 'gdef', 'tb'),
  Social = c('n', 'nf', 'shp', 'shw'),
  Prices = c('pa', 'pid', 'pg', 'pim'),
  Employment = c('n', 'nf'),
  Inequality = c('ydw', 'ydc', 'ineq', 'shp', 'shw'),
  Debt.and.Wealth = c('lh', 'lf', 'v', 'k', 'b_s'),
  Ecological = c('x_mat', 'mat', 'rec', 'emis', 'wa')
)

# Load Baseline - Shock Data
df <- shock.long.new(baseline, ple.shock)

# Plot Selected List Variables (scaled only if gdef is not there)
for (i in 1:length(selected.list)) {
  if ('gdef' %in% selected.list[[i]]) {
    print(view.vars(
      data = df,
      viz.vars = selected.list[[i]],
      var.label = names(selected.list)[i]
    ))
  } else {
    print(view.scaled.vars(
      data = df,
      viz.vars = selected.list[[i]],
      var.label = names(selected.list)[i]
    ))
  }
}

# Print Some Scaled Variables Example
# view.scaled.vars(data = shock.long.new(baseline, ple.shock),
#                  viz.vars = c('cab', 'ydw', 'n', 'rva'),
#                  var.label = 'Scaled')

# Print Some Variables Example (not scaled)
# view.vars(data = shock.long.new(baseline, ple.shock),
#           viz.vars = c('cab', 'ydw', 'n', 'rva'),
#           var.label = 'Not Scaled')
```

## Visualize Selected Variables Without Function

```{r shock.visualize.selected.2}
# Selected Vars to plot
selected <- c('n', 'c', 'ineq', 'va', 'cab', 'nf', 'gdef', 'tb', 'go')

p.selected.vars <- filter(df, variable %in% selected) %>%
  ggplot(aes(x = time, y = value, color = area, linetype = scenario)) +
  facet_wrap(~name, scales = 'free') +
  geom_line() +
  geom_vline(
    xintercept = initial$pars['t.shock', 'value'],
    linetype = 'dashed',
    linewidth = .4
  ) +
  labs(
    title = shock.title,
    subtitle = 'Selected Indicators. Vertical dashed line indicates shock time'
  )

p.selected.vars
```


## Summary Table

```{r ple.summary}
# Summary Table for PLE Scenario
ple.table <- shock.summary(
  baseline,
  shock.run = ple.shock,
  t_ = c(t.shock + 1, t.shock + 7, t.shock + 20),
  t_names = c('Immediate', 'Short.Term', 'Long.Term')
)

kable(ple.table, booktabs = T)
```

# Shock Example 4: Higher Use of Circular Economy Inputs

## Run Shock

This is a simple visualization of shock 4. Please note the new grammar to change initial values.

```{r shock.run.2}
# Shock Selection and Run
init.ce <- load.new.init(
  'data/perfect_init_aug14_withBmatrix_CE_inputs_higher_costs.xlsx'
)

# Time Period Selection
t.shock <- init.ce$pars['t.shock', 'value'] # shock time period
t0 <- 55 # initial time period for visualization
tf <- init.ce$pars['nPeriods', 'value'] # final time period for visualization

init.ce$pars['shock', 'value'] <- 4

# ce.shock <- run.new.model(init.ce, mvp.model)
# saveRDS(ce.shock, file = 'data/CE_shock_aug14.RDS')
ce.shock <- readRDS(file = 'data/CE_shock_aug14.RDS')

shock.title <- 'Scenario: Higher Use of Circular Economy Inputs'
```

## Visualize Shocked Variables

```{r shock.view.2}
view.shock(
  ce.shock,
  shock.vars = c('delta', 'vat-12'),
  shock.sectors = c(NA, 12)
)
```

## Visualize Impacts

```{r shock.visualize.scaled.2}
# Variable Selection
selected.list <- list(
  Economic = c('n', 'c', 'ineq', 'va', 'cab', 'nf', 'gdef', 'tb', 'go'),
  Macroeconomic = c('c', 'rva', 'go', 'id'),
  External = c('cab', 'gdef', 'tb'),
  Social = c('n', 'nf', 'shp', 'shw'),
  Prices = c('pa', 'pid', 'pg', 'pim'),
  Employment = c('n', 'nf'),
  Inequality = c('ydw', 'ydc', 'ineq', 'shp', 'shw'),
  Debt.and.Wealth = c('lh', 'lf', 'v', 'k', 'b_s'),
  Ecological = c('x_mat', 'mat', 'rec', 'emis', 'wa')
)

# Load Baseline - Shock Data
df <- shock.long.new(baseline, ce.shock)

# Plot Selected List Variables (scaled only if gdef is not there)
for (i in 1:length(selected.list)) {
  if ('gdef' %in% selected.list[[i]]) {
    print(view.vars(
      data = df,
      viz.vars = selected.list[[i]],
      var.label = names(selected.list)[i]
    ))
  } else {
    print(view.scaled.vars(
      data = df,
      viz.vars = selected.list[[i]],
      var.label = names(selected.list)[i]
    ))
  }
}

# Print Some Scaled Variables Example
# view.scaled.vars(data = shock.long.new(baseline, ple.shock),
#                  viz.vars = c('cab', 'ydw', 'n', 'rva'),
#                  var.label = 'Scaled')

# Print Some Variables Example (not scaled)
# view.vars(data = shock.long.new(baseline, ple.shock),
#           viz.vars = c('cab', 'ydw', 'n', 'rva'),
#           var.label = 'Not Scaled')
```

## Summary Table

```{r summary.table.2}
ce.table <- shock.summary(
  baseline,
  shock.run = ce.shock,
  t_ = c(t.shock + 1, t.shock + 7, t.shock + 20),
  t_names = c('Immediate', 'Short.Term', 'Long.Term')
)

kable(ce.table, booktabs = T)
```

# Appendix: Table of Industries

```{r industry.table}
kable(
  data.frame(number = 1:length(initial$sectors), industry = initial$sectors),
  booktabs = T
)
```